---
output: 
  word_document:
    reference_docx: T:/Temperature_TMDL_Revisions/model_QAPPs/R/data/rmd_template.docx
    keep_md: yes
    toc: yes
    fig_caption: yes
    fig_width: 12
    fig_height: 6
---

```{r setup_chap6x6, include=FALSE}

knitr::opts_chunk$set(results = "asis",
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      autodep = TRUE)

options(kableExtra.auto_format = F)

# For testing: model_extent <- "Model 2: Lemolo Powerhouse #1 to Toketee Reservoir"

```

##### `r model_extent`

```{r, label=t6x6j, include=FALSE}

t6x6j_met <- section.model.input %>% 
  dplyr::filter(`Model Extent` %in%  model_extent) %>% 
  dplyr::filter(`Model Location Type` %in%  c("Meteorological")) %>% 
  dplyr::mutate(`Data Source` = ifelse(`Station ID` %in% "TIR", strip_alpha(`Data Source`), `Data Source`),
                `Data Source` = ifelse(`Station ID` %in% "TIR", paste0(gsub(",.*$", "", `Data Source`)," (",stringi::stri_sub(`Data Source`,-4),")"), `Data Source`),
                `Data Source` = ifelse(is.na(`Data Source`),"",`Data Source`),
                `Data Source` = ifelse(is.na(`Interpolated Data`), `Data Source`,paste0("Derived data. ",`Data Source`))) %>% 
  dplyr::group_by(`Station ID`,`Model Location Type`, `Location Units`, `Data Source`,Note) %>% 
  dplyr::summarize(`Model Location` = toString(unique(sort(`Model Location`))),
                   Parameter = toString(unique(sort(Parameter))))%>% 
  dplyr::ungroup() %>% 
  dplyr::rename(`Model Location Name (Station ID)` = `Station ID`,
                `Input Type` = `Model Location Type`)

t6x6j_1 <- section.model.input %>% 
  dplyr::filter(`Model Extent` %in%  model_extent) %>% 
  dplyr::filter(!`Model Location Type` %in%  c("Calibration Site", "Meteorological")) %>% 
  dplyr::mutate(`Model Location Name (Station ID)` = ifelse(`Station ID` == "No Station ID" | is.na(`Station ID`),
                                                            `Model Location Name`,
                                                            paste0(`Model Location Name`, " (", `Station ID`, ")")),
                `Model Location` = round(`Model Location`,2),
                `Data Source` = ifelse(`Station ID` %in% "TIR", strip_alpha(`Data Source`), `Data Source`),
                `Data Source` = ifelse(`Station ID` %in% "TIR", paste0(gsub(",.*$", "", `Data Source`)," (",stringi::stri_sub(`Data Source`,-4),")"), `Data Source`),
                `Data Source` = ifelse(is.na(`Data Source`),"",`Data Source`),
                `Data Source` = ifelse(is.na(`Interpolated Data`), `Data Source`, paste0("Derived data. ",`Data Source`))) %>%   
  dplyr::select(`Model Location Name (Station ID)`, `Model Location`, `Location Units`, `Model Location Type`, `Parameter`, `Data Source`, Note) %>%
  dplyr::rename(`Input Type` = `Model Location Type`) %>% 
  rbind(t6x6j_met) %>% 
  dplyr::arrange(`Input Type`,Parameter,desc(`Model Location`))


t6x6j_col <- t6x6j_1 %>% 
  dplyr::filter(!is.na(`Location Units`))

names(t6x6j_1)[names(t6x6j_1) == "Model Location"] <- if(NROW(t6x6j_col)>0){paste0("Model Location (",unique(t6x6j_col$`Location Units`),")")}else{paste0("Model Location")}

t6x6j_1 <- t6x6j_1[,-c(3)]%>% 
    dplyr::rename(`Model Input` = Parameter)

if(NROW(t6x6j_1)>0){
  
  t6x6j_tbl <- knitr::kable(t6x6j_1, format = "pandoc", padding = 2,
                            caption = tbls(name = paste0("t6x6j_1", model_extent),
                                           caption = paste0("Boundary condition and tributary inputs to the existing ", model_extent, ".")))
  
}

```
`r if(NROW(t6x6j_1)>0){paste0(tbls(name = paste0("t6x6j_1", model_extent),display="cite")," summarizes the current configuration of the model input parameters and the source of these data. Temperature, flow, and meteorological input parameters are summarized to improve documentation of the TMDL approach.")}`

`r if(NROW(t6x6j_1)>0){t6x6j_tbl}else{paste0("There were no boundary condition and tributary inputs to this model.")}`