---
output: 
  word_document:
    reference_docx: T:/Temperature_TMDL_Revisions/model_QAPPs/R/data/rmd_template.docx
    keep_md: yes
    toc: yes
    fig_caption: yes
    fig_width: 12
    fig_height: 6
---

```{r setup_chap6x7, include=FALSE}

knitr::opts_chunk$set(results = "asis",
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      autodep = TRUE)

options(kableExtra.auto_format = F)
options(knitr.kable.NA = '')

# For testing: model_extent <- "Model 2: Lemolo Powerhouse #1 to Toketee Reservoir"

```

#### `r model_extent`

```{r, label=`t6x7j`, include=FALSE}

t6x7j <- section.model.input %>% 
  dplyr::filter(`Model Extent` %in%  model_extent) %>% 
  dplyr::filter(`Model Location Type` %in% "Calibration Site") %>%
  dplyr::mutate(`Data Source` = ifelse(`Station ID` %in% "TIR", strip_alpha(`Data Source`), `Data Source`),
                `Data Source` = ifelse(`Station ID` %in% "TIR", paste0(gsub(",.*$", "", `Data Source`)," (",stringi::stri_sub(`Data Source`,-4),")"), `Data Source`),
                `Data Source` = ifelse(is.na(`Data Source`),"",`Data Source`),
                `Data Source` = ifelse(is.na(`Interpolated Data`), `Data Source`, paste0("Derived data. ",`Data Source`)),
                Parameter = ifelse(`Station ID` %in% "TIR", paste0(Parameter," (TIR)"), Parameter),
                `Model Location` = round(as.numeric(`Model Location`),2),
                `Model Location` = ifelse(`Station ID` %in% "TIR","Model extent",`Model Location`),
                `Model Location Name` = ifelse(`Station ID` %in% "TIR","Model extent",`Model Location Name`),
                `Station ID` = ifelse(`Station ID` %in% "TIR",NA,`Station ID`),
                `Model Location Name (Station ID)` = ifelse(`Station ID` == "No Station ID" | is.na(`Station ID`),
                                                            `Model Location Name`,
                                                            paste0(`Model Location Name`, " (",`Station ID`,")")))%>% 
  dplyr::group_by(`Model Location Name (Station ID)`,`Model Location`,`Location Units`,`Data Source`) %>% 
  dplyr::summarize(Parameter = toString(unique(sort(Parameter)))) %>% 
  dplyr::ungroup() %>% 
  dplyr::arrange(Parameter,desc(as.numeric(`Model Location`))) %>% 
  dplyr::rename(`Calibration Parameter` = `Parameter`)

t6x7j_col <- t6x7j %>% 
  dplyr::filter(!is.na(`Location Units`))

names(t6x7j)[names(t6x7j) == "Model Location"] <- if(NROW(t6x7j_col)>0){paste0("Model Location (",unique(t6x7j_col$`Location Units`),")")}else{paste0("Model Location")}

t6x7j <- t6x7j[,c(1,2,5,4)]

if(NROW(t6x7j)>0){
  
  t6x7j_tbl <- knitr::kable(t6x7j, format = "pandoc", padding = 2,
                            caption = tbls(name = paste0("t6x7j", model_extent, waterbody_name),
                                           caption = paste0("Calibration sites and parameters used in the ", model_extent, " Heat Source Model.")))
  
  mod.loc <- ifelse(tmdl.mod.2[which(tmdl.mod.2$`Model Waterbody`== waterbody_name),]$mod_rmd %in% c("hs7","hs8","hs9"),
                    print("The model location in the table below describes the distance of each input from the most downstream model node."),
                    ifelse(tmdl.mod.2[which(tmdl.mod.2$`Model Waterbody`== waterbody_name),]$mod_rmd %in% c("hs6"),
                           print("The model location in the table below describes the distance of each input from the most upstream model node."),
                           print("")))
  
}

```

`r if(NROW(t6x7j)>0){paste0("The expected model calibration sites and data sources are summarized in ",tbls(name = paste0("t6x7j", model_extent, waterbody_name),display="cite"),". The model location in the table below describes the distance of each input from the most downstream model node. The model inputs and parameters that are expected to be modified to improve model fit are described in Section 6.1.")}`

`r if(NROW(t6x7j)>0){t6x7j_tbl}`

`r if(waterbody_name == "North Umpqua River (Spawning period)" & model_extent == "Model 4: Slide Powerhouse to Soda Springs Reservoir"){paste0("There are no calibration sites available for the North Umpqua River (Spawning period) Model 4: Slide Powerhouse to Soda Springs Reservoir.")}`
