---
output: 
  word_document:
    reference_docx: T:/Temperature_TMDL_Revisions/model_QAPPs/R/data/rmd_template.docx
    keep_md: yes
    toc: yes
    fig_caption: yes
    fig_width: 12
    fig_height: 6
---

```{r, label=`setup_chap6x6`, include=FALSE}

knitr::opts_chunk$set(results = "asis",
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      autodep = TRUE)

options(kableExtra.auto_format = F)

# For testing: model_extent <- "Model 2: Lemolo Powerhouse #1 to Toketee Reservoir"

```

#### `r model_extent`

```{r, label=`t6x6j`, include=FALSE}

t6x6j_met <- section.model.input %>% 
  dplyr::filter(`Model Extent` %in%  model_extent) %>% 
  dplyr::filter(`Model Location Type` %in%  c("Meteorological")) %>% 
  dplyr::mutate(`Data Source` = ifelse(substr(`Data Source`,1,18) == "Watershed Sciences", 
                                       strip_alpha(`Data Source`), `Data Source`),
                `Data Source` = ifelse(substr(`Data Source`,1,18) == "Watershed Sciences", 
                                       paste0(gsub(",.*$", "", `Data Source`)," (",stringi::stri_sub(`Data Source`,-4),")"), `Data Source`),
                `Data Source` = ifelse(is.na(`Data Source`),"",`Data Source`),
                `Data Source` = ifelse(is.na(`Interpolated Data`), `Data Source`,paste0("Derived data. ",ifelse(`Data Source`=="DEQ","",`Data Source`)))) %>% 
  dplyr::group_by(`Station ID`,`Model Location Type`, `Location Units`, `Data Source`,Note) %>% 
  dplyr::summarize(`Model Location` = toString(unique(sort(round(as.numeric(`Model Location`),2)))),
                   Parameter = toString(unique(sort(Parameter))))%>% 
  dplyr::ungroup() %>% 
  dplyr::rename(`Model Location Name (Station ID)` = `Station ID`,
                `Input Type` = `Model Location Type`)

t6x6j_1 <- section.model.input %>% 
  dplyr::filter(`Model Extent` %in%  model_extent) %>% 
  dplyr::filter(!`Model Location Type` %in%  c("Calibration Site", "Meteorological")) %>% 
  dplyr::mutate(`Model Location Name (Station ID)` = ifelse(`Station ID` %in%  "No Station ID" | is.na(`Station ID`),
                                                            `Model Location Name`,
                                                            paste0(`Model Location Name`, " (", `Station ID`, ")")),
                `Model Location` = round(as.numeric(`Model Location`),2),
                `Data Source` = ifelse(substr(`Data Source`,1,18) == "Watershed Sciences", 
                                       strip_alpha(`Data Source`), `Data Source`),
                `Data Source` = ifelse(substr(`Data Source`,1,18) == "Watershed Sciences", 
                                       paste0(gsub(",.*$", "", `Data Source`)," (",stringi::stri_sub(`Data Source`,-4),")"), `Data Source`),
                `Data Source` = ifelse(is.na(`Data Source`),"",`Data Source`),
                `Data Source` = ifelse(is.na(`Interpolated Data`), `Data Source`, paste0("Derived data. ",ifelse(`Data Source`=="DEQ","",`Data Source`)))) %>%   
  dplyr::select(`Model Location Name (Station ID)`, `Model Location`, `Location Units`, `Model Location Type`, `Parameter`, `Data Source`, Note) %>%
  dplyr::rename(`Input Type` = `Model Location Type`) %>% 
  dplyr::arrange(`Input Type`,Parameter,desc(as.numeric(`Model Location`))) %>% 
  rbind(t6x6j_met)

t6x6j_col <- t6x6j_1 %>% 
  dplyr::filter(!is.na(`Location Units`))

names(t6x6j_1)[names(t6x6j_1) == "Model Location"] <- if(NROW(t6x6j_col)>0){paste0("Model Location (",unique(t6x6j_col$`Location Units`),")")}else{paste0("Model Location")}

t6x6j_1 <- t6x6j_1[,-c(3)]%>% 
  dplyr::rename(`Model Input` = Parameter)

if(NROW(t6x6j_1)>0){
  
  t6x6j_tbl <- knitr::kable(t6x6j_1, format = "pandoc", padding = 2,
                            caption = tbls(name = paste0("t6x6j_1", model_extent, waterbody_name),
                                           caption = paste0("Boundary condition and tributary inputs to the ", model_extent, ".")))
  
}

```
`r if(NROW(t6x6j_1)>0){paste0(tbls(name = paste0("t6x6j_1", model_extent, waterbody_name),display="cite")," summarizes the current configuration of the model input parameters and the source of these data. Temperature, flow, and meteorological input parameters are summarized to improve documentation of the TMDL approach.")}`

`r if(NROW(t6x6j_1)>0){t6x6j_tbl}else{paste0("There were no boundary condition and tributary inputs to this model.")}`

```{r, label=`t6x6j_2`, echo=FALSE}

t6x6j_2 <- section.model.input %>%
  dplyr::filter(`Model Extent` %in%  model_extent) %>% 
  dplyr::filter(`Model Location Type` %in% "Meteorological")

```

`r if(NROW(t6x6j_2)>1){paste0("Hourly meteorology inputs into the model include ", tolower(knitr::combine_words(unique(sort(t6x6j_2$Parameter)))),". Air temperature data were modified using the dry adiabatic lapse rate to adjust for differences in elevation between the measurement location and the model input location. Wind speeds were adjusted to improve the calibration using a wind-sheltering coefficient to represent difference in wind speed between the measurement location and above the stream within the riparian area.")}else{paste0("There are no meteorology inputs into the model.")}`

`r if(!is.na(unique(section.model.info$"Veg Notes"))){paste0("The near-stream vegetation inputs to the model include vegetation height and canopy cover. ", unique(section.model.info$"Veg Notes"))}`
