---
output: 
  word_document:
    reference_docx: T:/Temperature_TMDL_Revisions/model_QAPPs/R/data/rmd_template.docx
    keep_md: yes
    toc: yes
    fig_caption: yes
    fig_width: 12
    fig_height: 6
---

```{r, label=`setup-chap6`, include=FALSE}
knitr::opts_chunk$set(results = "asis",
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      autodep = TRUE)
options(kableExtra.auto_format = F)
options(knitr.kable.NA = '')

# for test: waterbody_name <- "North Umpqua River"

section.model.info <- model.info %>% 
  dplyr::filter(`Model Waterbody` %in%  waterbody_name)

section.model.input <- model.input %>% 
  dplyr::filter(`Model Waterbody` %in%  waterbody_name)

model.extent <- ifelse(NROW(section.model.input)>0,
                       ifelse(section.model.input$`Model Waterbody` == section.model.input$`Model Extent`, TRUE, FALSE),
                       TRUE)

model.extent <- na.omit(model.extent)

```

## `r waterbody_name`

`r paste0("The ",gsub("SHADOW ","",waterbody_name)," SHADOW models are shade models developed by the ",unique(section.model.info[which(section.model.info$"Model Waterbody" == waterbody_name),]$"Calibration Org"),". DEQ does not intend to update these models. If the need arises to revisit some of the modeling, this QAPP will be updated as needed.")`

### Model domain

```{r, label=`t6x1`, include=FALSE}

model.domain <- unique(sort(section.model.info$`Model Extent Text`))

```

The extent of the model domain is `r model.domain` The model extent is shown in the HTML interactive map that accompanies this QAPP and is referenced in Appendix C.

