---
output: 
  word_document:
    reference_docx: T:/Temperature_TMDL_Revisions/model_QAPPs/R/data/rmd_template.docx
    keep_md: yes
    toc: yes
    fig_caption: yes
    fig_width: 12
    fig_height: 6
---

```{r, label=`setup-chap6`, include=FALSE}
knitr::opts_chunk$set(results = "asis",
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      autodep = TRUE)
options(kableExtra.auto_format = F)
options(knitr.kable.NA = '')

# for test: waterbody_name <- "North Umpqua River"

section.model.info <- model.info %>% 
  dplyr::filter(`Model Waterbody` %in%  waterbody_name)

section.model.input <- model.input %>% 
  dplyr::filter(`Model Waterbody` %in%  waterbody_name)

model.extent <- ifelse(NROW(section.model.input)>0,
                       ifelse(section.model.input$`Model Waterbody` == section.model.input$`Model Extent`, TRUE, FALSE),
                       TRUE)

model.extent <- na.omit(model.extent)

```

## `r waterbody_name`

The `r waterbody_name` model is a `r tolower(ifelse(unique(section.model.info[which(section.model.info$"Model Waterbody" == waterbody_name),]$"Primary Model Parameter") == "Solar", "shade",unique(section.model.info[which(section.model.info$"Model Waterbody" == waterbody_name),]$"Primary Model Parameter")))` model developed using `r unique(section.model.info[which(section.model.info$"Model Waterbody" == waterbody_name),]$"Model version")`. The model `r if(section.model.info[which(section.model.info$"Model Waterbody" == waterbody_name),]$"New Model" %in% "T"){paste0("will be")}else{paste0("was")}` developed by `r unique(section.model.info[which(section.model.info$"Model Waterbody" == waterbody_name),]$"Calibration Org")`.

### Model domain

```{r, label=`t6x1`, include=FALSE}

if(length(section.model.info$`Model Extent Text`)>1){
  
  model.domain <- unique(sort(section.model.info$`Model Extent Text`))
  
} else {
  
  model.domain <- section.model.info$`Model Extent Text`
  
} 
```

`r if(waterbody_name == "Southern Willamette Subbasins"){model.domain}else{paste0("The extent of the model domain is the ", model.domain)}` The model extent is shown in the HTML interactive map that accompanies this QAPP and is referenced in Appendix C.

### Spatial and temporal resolution

The model input spatial resolution (*dx*) is `r unique(section.model.info$"Input dx")` `r unique(section.model.info$"Input dx Units")`. Outputs are generated every `r unique(section.model.info$"Output dx")` `r unique(section.model.info$"Output dx Units")`. The model time step (*dt*) is `r unique(section.model.info$dT)` `r unique(section.model.info$"dT Units")` and outputs are generated every hour.

A *dx* of `r unique(section.model.info$"Input dx")` `r unique(section.model.info$"Input dx Units")` was chosen to capture the range of solar flux input caused by the varied vegetation conditions along the length of the stream. The high resolution *dx* will allow evaluation of multiple vegetation management scenarios for each designated management agency.

### Source characteristics

```{r, label=`sc`, include=FALSE}

## 1st paragraph
source.char.1 <- readxl::read_xlsx("//deqhq1/TMDL/Planning statewide/Temperature_TMDL_Revisions/model_QAPPs/R/data/tables_SouthWillamette.xlsx",
                                 sheet = "sW_sourceChar") %>%
  dplyr::filter(`Model Waterbody` == waterbody_name) %>% 
  dplyr::pull(sourceChar1)

source.char.2 <- readxl::read_xlsx("//deqhq1/TMDL/Planning statewide/Temperature_TMDL_Revisions/model_QAPPs/R/data/tables_SouthWillamette.xlsx",
                                 sheet = "sW_sourceChar") %>%
  dplyr::filter(`Model Waterbody` == waterbody_name) %>% 
  dplyr::pull(sourceChar2)

# NPDES
npdes.ind.waterbody <- npdes.ind.pro.area %>%
  dplyr::filter(grepl(pattern=gsub("\\s*\\([^\\)]+\\)","",x=waterbody_name), x=`Stream Name`, ignore.case = TRUE)) %>% 
  dplyr::mutate(`Facility Name (Facility Number)` = paste0(`Common Name`," (", `WQ File Nbr`,")"),
                `Latitude/Longitude` = paste0(round(Latitude,4), "/",round(Longitude,3)),
                `Permit Type and Description` = paste0(`Permit Type`, ": ", `Permit Description`),
                `River Mile` = round(`River Mile`,1),
                `Stream/River Mile` = ifelse(is.na(`Stream Name`), NA, paste0(`Stream Name`, " ", " RM ",`River Mile`))) %>% 
  dplyr::select(`Facility Name (Facility Number)`, `Latitude/Longitude`,`Permit Type and Description`, `Stream/River Mile`) %>% 
  dplyr::arrange(`Facility Name (Facility Number)`)

npdes.ind.n <- nrow(npdes.ind.waterbody)

if(npdes.ind.n>0){
  
  npdes.ind.tbl <- knitr::kable(npdes.ind.waterbody, format = "pandoc", padding = 2,
                                caption = tbls(name = paste0("npdes.ind.waterbody_",waterbody_name),
                                               caption = ifelse(waterbody_name == "McKenzie River: Upper", 
                                                                paste0("Summary of individual NPDES permitted discharges in the upper portion of the McKenzie River."),
                                                                paste0("Summary of individual NPDES permitted discharges in the ", waterbody_name, "."))))
  
}

if(npdes.ind.n == 0) {
  npdes.txt <- "There are no permitted individual NPDES point sources along the model extent."
} 

if(npdes.ind.n == 1) {
  npdes.txt <- paste0("There is one permitted individual NPDES point source along the model extent. Detail about the point source is summarized in ",tbls(name = paste0("npdes.ind.waterbody_",waterbody_name),display="cite"), ".")
}

if(npdes.ind.n > 1) {
  npdes.txt <- paste0("There are ", numbers.to.words(npdes.ind.n)," permitted individual NPDES point sources along the model extent. Detail about each point source is summarized in ",tbls(name = paste0("npdes.ind.waterbody_",waterbody_name),display="cite"),".")
} 

# NLCD
nlcd.waterbody <- nlcd.pro.area %>%
  dplyr::mutate(Stream = ifelse(Stream == "McKenzie River", "McKenzie River: Upper", Stream)) %>% # Temporary edits - check
  dplyr::filter(Stream == waterbody_name) %>% 
  dplyr::select(NLCD_Land,Acres,Percentage) %>% 
  dplyr::arrange(desc(Acres)) %>% 
  dplyr::rename("2016 NLCD Land Cover" = NLCD_Land,
                "Percent of Total Acres" = Percentage)

if(NROW(nlcd.waterbody)>0){
  
  nlcd.wb.tbl <- knitr::kable(nlcd.waterbody, format = "pandoc", padding = 2,
                              caption = tbls(name = paste0("nlcd.waterbody_",waterbody_name),
                                             caption = ifelse(waterbody_name == "McKenzie River: Upper", 
                                                                paste0("Summary of land uses within 100 meters of the digitized centerline of the upper portion of the McKenzie River based on the 2016 National Land Cover Database (Yang et al 2018)."),
                                                                paste0("Summary of land uses within 100 meters of the digitized ", waterbody_name, " centerline based on the 2016 National Land Cover Database (Yang et al 2018)."))))
  
}

nlcd.text.waterbody <- nlcd.text.pro.area %>% 
  dplyr::mutate(Stream = ifelse(Stream == "McKenzie River", "McKenzie River: Upper", Stream)) %>% # Temporary edits - check
  dplyr::filter(Stream == waterbody_name)

# DMA
dma.waterbody <- dma.pro.area %>% 
  dplyr::mutate(Stream = ifelse(Stream == "McKenzie River", "McKenzie River: Upper", Stream)) %>% # Temporary edits - check
  dplyr::filter(Stream == waterbody_name) %>% 
  dplyr::select(DMA_RP,Acres,Percentage) %>% 
  dplyr::arrange(desc(Acres)) %>% 
  dplyr::rename("DMA or Responsible Person" = DMA_RP,
                "Percent of Total Acres" = Percentage) 

if(NROW(dma.waterbody)>0){
  
  dma.wb.tbl <- knitr::kable(dma.waterbody, format = "pandoc", padding = 2,
                             caption = tbls(name = paste0("dma.waterbody",waterbody_name),
                                            caption = ifelse(waterbody_name == "McKenzie River: Upper", 
                                                                paste0("Summary of designated management agencies (DMAs) or responsible persons within 100 meters of the digitized centerline of the upper portion of the McKenzie River."),
                                                                paste0("Summary of designated management agencies (DMAs) or responsible persons within 100 meters of the digitized ", waterbody_name, " centerline."))))
  
}

```

The primary sources of thermal loading contributing to temperatures exceedances along the `r if(waterbody_name == "McKenzie River: Upper"){paste0("upper portion of the McKenzie River")}else{waterbody_name}` include increases in solar radiation loading from the disturbance or removal of near-stream vegetation, `r if(npdes.ind.n>0){paste0("point source discharges,")}` `r if(!is.na(source.char.1)){source.char.1}` and background sources (`r strip_alpha(section.model.info$"Abbreviated Reference")`). `r source.char.2` The contribution of these latter potential sources may be investigated as part of the model scenarios.

`r npdes.txt`

`r if(npdes.ind.n > 0){npdes.ind.tbl}`

The majority land use`r if(NROW(nlcd.text.waterbody)>0){s(nlcd.text.waterbody$n)}` along the `r if(waterbody_name == "McKenzie River: Upper"){paste0("upper portion of the McKenzie River")}else{waterbody_name}` `r if(NROW(nlcd.text.waterbody)>0){is.are(nlcd.text.waterbody$n[1])}` `r if(NROW(nlcd.text.waterbody)>0){nlcd.text.waterbody$text}` accounting for about `r if(NROW(nlcd.text.waterbody)>0){nlcd.text.waterbody$MajorityPercent}` percent of the near stream area. `r if(NROW(nlcd.waterbody)>0){tbls(name = paste0("nlcd.waterbody_",waterbody_name),display="cite")}` summarizes all the land uses within 100 meters of the digitized `r if(waterbody_name == "McKenzie River: Upper"){paste0("upper portion of the McKenzie River")}else{waterbody_name}` centerline. Land uses were summarized using the 2016 National Land Cover Database (Yang et al 2018). Note that Shrub/Scrub and Herbaceous land uses can be areas where forest clearcuts have occurred and would be classified as forest after regrowth.

`r if(NROW(nlcd.waterbody)>0){nlcd.wb.tbl}`

Anthropogenic related stream warming caused by nonpoint sources is closely associated with the uses, activity, and the condition of vegetation adjacent to the stream. How activities and uses are managed in these areas is partially determined by a variety of different rules and management plans established by the landowner and any agency with land use authority. To better understand the spatial distribution of different agency rules or management plans along the model extent DEQ mapped known designated management agencies.

A designated management agency is defined in OAR 340-042-0030(2) as a federal, state, or local governmental agency that has legal authority over a sector or source contributing pollutants. Typically, persons or designated management agencies that are identified in the TMDL Water Quality Management Plan (WQMP) are responsible for developing TMDL implementation plans and implementing management strategies to reduce pollutant loading. `r if(NROW(dma.waterbody)>0){tbls(name = paste0("dma.waterbody",waterbody_name),display="cite")}` summarizes the potential designated management agencies and responsible persons along the `r if(waterbody_name == "McKenzie River: Upper"){paste0("upper portion of the McKenzie River")}else{waterbody_name}` model extent. 

`r if(NROW(dma.waterbody)>0){dma.wb.tbl}`

### Time frame of simulation

```{r, label=`tf`, include=FALSE}

tf <- tmdl.mod.2 %>% 
  dplyr::filter(`Model Waterbody` == waterbody_name) %>% 
  pull(Model_version)

```

`r if(tf == "Heat Source Version 6"){paste0("The model period is for a single day: ",format(as.Date(as.numeric(unique(section.model.info$"Model Period Start")), origin = "1899-12-30"), format="%B %d, %Y"),".")}else{paste0("The model period is ", format(as.Date(as.numeric(unique(section.model.info$"Model Period Start")), origin = "1899-12-30"), format="%B %d, %Y"), " to ", format(as.Date(as.numeric(unique(section.model.info$"Model Period End")), origin = "1899-12-30"), format="%B %d, %Y"), ".")}`

### Important assumptions

The effort currently described in the QAPP includes use of existing models. Key calibration assumptions made during the model setup and calibration process were documented in the original TMDL `r paste0("(", knitr::combine_words(strip_alpha(unique(sort(model.info[is.na(model.info$'Reference Report'),]$"Abbreviated Reference")))), ")")`, the model user guide `r knitr::combine_words(strip_alpha(unique(sort(unique(tmdl.mod.2$mod_ref_intext)))))` or in Section 6 and Section 7 of this document.

### Model inputs

```{r, label=`t6x6_1`, include=FALSE}
options(knitr.kable.NA = '') # hide NAs in table

if(!model.extent == TRUE){
  
  chap6x6 <- NULL
  
  for(model_extent in unique(sort(section.model.input$`Model Extent`))){
    
    chap6x6 <- c(chap6x6, knitr::knit_child(input = "chap6x6.Rmd", envir = globalenv()))
    
  }
  
} else {
  
  t6x6_met <- section.model.input %>% 
    dplyr::filter(`Model Location Type` %in%  c("Meteorological")) %>%
    dplyr::mutate(`Data Source` = ifelse(`Station ID` %in% "TIR", strip_alpha(`Data Source`), `Data Source`),
                  `Data Source` = ifelse(`Station ID` %in% "TIR", paste0(gsub(",.*$", "", `Data Source`)," (",stringi::stri_sub(`Data Source`,-4),")"), `Data Source`),
                  `Data Source` = ifelse(is.na(`Data Source`),"",`Data Source`),
                  `Data Source` = ifelse(is.na(`Interpolated Data`), `Data Source`,paste0("Derived data. ",`Data Source`))) %>% 
    dplyr::group_by(`Station ID`,`Model Location Type`, `Location Units`, `Data Source`,Note) %>% 
    dplyr::summarize(`Model Location` = toString(unique(sort(round(`Model Location`,2)))),
                     Parameter = toString(unique(sort(Parameter))))%>% 
    dplyr::ungroup() %>%
    dplyr::rename(`Model Location Name (Station ID)` = `Station ID`,
                  `Input Type` = `Model Location Type`)
  
  t6x6_1 <- section.model.input %>% 
    dplyr::filter(!`Model Location Type` %in%  c("Calibration Site", "Meteorological")) %>% 
    dplyr::mutate(`Model Location Name (Station ID)` = ifelse(`Station ID` %in% "No Station ID" | is.na(`Station ID`),
                                                              `Model Location Name`,
                                                              paste0(`Model Location Name`, " (", `Station ID`, ")")),
                  `Model Location` = round(`Model Location`,2),
                  `Data Source` = ifelse(`Station ID` %in% "TIR", strip_alpha(`Data Source`), `Data Source`),
                  `Data Source` = ifelse(`Station ID` %in% "TIR", paste0(gsub(",.*$", "", `Data Source`)," (",stringi::stri_sub(`Data Source`,-4),")"), `Data Source`),
                  `Data Source` = ifelse(is.na(`Data Source`),"",`Data Source`),
                  `Data Source` = ifelse(is.na(`Interpolated Data`),`Data Source`, paste0("Derived data. ",`Data Source`))) %>%  
    dplyr::select(`Model Location Name (Station ID)`, `Model Location`, `Location Units`, `Model Location Type`, `Parameter`, `Data Source`, Note) %>%
    dplyr::rename(`Input Type` = `Model Location Type`) %>% 
    rbind(t6x6_met) %>% 
    dplyr::arrange(`Input Type`,Parameter,desc(`Model Location`))
  
  t6x6_col <- t6x6_1 %>% 
    dplyr::filter(!is.na(`Location Units`))
  
  names(t6x6_1)[names(t6x6_1) == "Model Location"] <- if(NROW(t6x6_col)>0){paste0("Model Location (",unique(t6x6_col$`Location Units`),")")}else{paste0("Model Location")}
  
  t6x6_1 <- t6x6_1[,-c(3)] %>% 
    dplyr::rename(`Model Input` = Parameter)
  
  if(NROW(t6x6_1)>0){
    
    t6x6_1_tbl <-  knitr::kable(t6x6_1, format = "pandoc", padding = 2,
                                caption = tbls(name = paste0("t6x6_1", waterbody_name),
                                               caption = ifelse(waterbody_name == "McKenzie River: Upper", 
                                                                paste0("Boundary condition and tributary inputs to the existing Heat Source Model for the upper portion of the McKenzie River."),
                                                                paste0("Boundary condition and tributary inputs to the existing ", waterbody_name, " Heat Source Model."))))
    
  }
  
  #t6x6_1.txt <- "**[Update for CE-QUAL-W2 model]**"
  
}

```

`r if(!model.extent == TRUE){} else {if(NROW(t6x6_1)>0){paste0(tbls(name = paste0("t6x6_1", waterbody_name),display="cite")," summarizes the current configuration of the model input parameters and the source of these data. Temperature, flow, and meteorological input parameters are summarized to improve documentation of the TMDL approach.")}}`

`r if(!model.extent == TRUE){paste(chap6x6, collapse = "\n")}else{if(NROW(t6x6_1)>0){t6x6_1_tbl}}`

```{r, label=`t6x6_2`, echo=FALSE}

t6x6_2 <- section.model.input %>%
  dplyr::filter(`Model Location Type` %in% "Meteorological")

```

`r if(NROW(t6x6_2)>1){paste0("Hourly meteorology inputs into the model include ", tolower(knitr::combine_words(unique(sort(t6x6_2$Parameter)))),". Air temperature data were modified using the dry adiabatic lapse rate to adjust for differences in elevation between the measurement location and the model input location. Wind speeds were adjusted using a wind-sheltering coefficient to represent difference in wind speed between the measurement location and above the stream within the riparian area.")}else{paste0("There are no meteorology inputs into the model.")}`

`r if(!is.na(unique(section.model.info$"Veg Notes"))){paste0("The near stream vegetation inputs to the model include vegetation height and canopy cover. ", unique(section.model.info$"Veg Notes"))}`

### Model calibration

```{r, label=`t6x7`, include=FALSE}
options(knitr.kable.NA = '') # hide NAs in table

if(!model.extent == TRUE){
  
  chap6x7 <- NULL
  
  for(model_extent in unique(sort(section.model.input$`Model Extent`))){
    
    chap6x7 <- c(chap6x7, knitr::knit_child(input = "chap6x7.Rmd", envir = globalenv()))
    
  }
  
} else {
  
  t6x7 <- section.model.input %>% 
    dplyr::filter(`Model Location Type` %in% "Calibration Site") %>%
    dplyr::mutate(`Data Source` = ifelse(`Station ID` %in% "TIR", strip_alpha(`Data Source`), `Data Source`),
                  `Data Source` = ifelse(`Station ID` %in% "TIR", paste0(gsub(",.*$", "", `Data Source`)," (",stringi::stri_sub(`Data Source`,-4),")"), `Data Source`),
                  `Data Source` = ifelse(is.na(`Data Source`),"",`Data Source`),
                  `Data Source` = ifelse(is.na(`Interpolated Data`), `Data Source`,paste0("Derived data. ",`Data Source`)),
                  Parameter = ifelse(`Station ID` %in% "TIR", paste0(Parameter," (TIR)"), Parameter),
                  `Model Location` = round(`Model Location`,2),
                  `Model Location` = ifelse(`Station ID` %in% "TIR","Model extent",`Model Location`),
                  `Model Location Name` = ifelse(`Station ID` %in% "TIR","Model extent",`Model Location Name`),
                  `Station ID` = ifelse(`Station ID` %in% "TIR",NA,`Station ID`),
                  `Model Location Name (Station ID)` = ifelse(`Station ID` == "No Station ID" | is.na(`Station ID`),
                                                              `Model Location Name`,
                                                              paste0(`Model Location Name`, " (",`Station ID`,")")))%>% 
    dplyr::group_by(`Model Location Name (Station ID)`,`Model Location`,`Location Units`,`Data Source`) %>% 
    dplyr::summarize(Parameter = toString(unique(sort(Parameter)))) %>% 
    dplyr::ungroup() %>% 
    dplyr::arrange(desc(as.numeric(`Model Location`))) %>% 
    dplyr::rename(`Calibration Parameter` = `Parameter`)
  
  t6x7_col <- t6x7 %>% 
    dplyr::filter(!is.na(`Location Units`))
  
  names(t6x7)[names(t6x7) == "Model Location"] <- if(NROW(t6x7_col)>0){paste0("Model Location (",unique(t6x7_col$`Location Units`),")")}else{paste0("Model Location")}
  
  t6x7 <- t6x7[,c(1,2,5,4)]
  
  if(NROW(t6x7)>0){
    
    if(waterbody_name == "Southern Willamette Subbasins"){
      
      # remove table column "Model Location"
      t6x7_tbl <- knitr::kable(t6x7[,-2], format = "pandoc", padding = 2,
                               caption = tbls(name = paste0("t6x7", waterbody_name),
                                              caption = paste0("Calibration sites and parameters used in the existing ", waterbody_name, " Heat Source Model.")))
      
    } else {
      
      t6x7_tbl <- knitr::kable(t6x7, format = "pandoc", padding = 2,
                               caption = tbls(name = paste0("t6x7", waterbody_name),
                                              caption = ifelse(waterbody_name == "McKenzie River: Upper", 
                                                                paste0("Calibration sites and parameters used in the existing Heat Source Model for the upper portion of the McKenzie River."),
                                                                paste0("Calibration sites and parameters used in the existing ", waterbody_name, " Heat Source Model."))))
      
    }
  }
  
  mod.loc <- ifelse(tmdl.mod.2[which(tmdl.mod.2$`Model Waterbody`== waterbody_name),]$mod_rmd %in% c("hs7","hs8","hs9"),
                    print("The model location in the table below describes the distance of each input from the most downstream model node."),
                    ifelse(tmdl.mod.2[which(tmdl.mod.2$`Model Waterbody`== waterbody_name),]$mod_rmd %in% c("hs6"),
                           print("The model location in the table below describes the distance of each input from the most upstream model node."),
                           print("")))
  
  t6x7_1.txt <- "**[Update for CE-QUAL-W2 model]**"
  
}

```

`r if(!waterbody_name == "Southern Willamette Subbasins"){if(!model.extent == TRUE){} else {if(NROW(t6x7)>0){paste0("Model calibration was completed by ", knitr::combine_words(unique(sort(section.model.info$"Calibration Org")))," (", knitr::combine_words(strip_alpha(unique(sort(section.model.info$"Abbreviated Reference")))),"). The model calibration sites and data sources are summarized in ", tbls(name = paste0("t6x7", waterbody_name),display="cite"), ". ", mod.loc[1], if(NROW(t56)>0){paste0(" Effective shade model calibrations sites are summarized in ",tbls(name = "t56",display="cite"),".")}," If it is determined that the model calibration needs to be updated, the model inputs and parameters that are expected to be modified to improve model fit are described in Section 6.1 General model inputs and parameters.")}else{t6x7_1.txt}}}else{paste0("Model calibration was completed by ", knitr::combine_words(unique(sort(section.model.info$"Calibration Org")))," (", knitr::combine_words(strip_alpha(unique(sort(section.model.info$"Abbreviated Reference")))),"). The model calibration sites and data sources are summarized in ", tbls(name = paste0("t6x7", waterbody_name),display="cite"), ". ", if(NROW(t56)>0){paste0("Effective shade model calibrations sites are summarized in ",tbls(name = "t56",display="cite"),".")}," If it is determined that the model calibration needs to be updated, the model inputs and parameters that are expected to be modified to improve model fit are described in Section 6.1 General model inputs and parameters.")}`

`r if(!model.extent == TRUE) {paste(chap6x7, collapse = "\n")} else {if(NROW(t6x7)>0){t6x7_tbl}}`
