---
output: 
  word_document:
    reference_docx: T:/Temperature_TMDL_Revisions/model_QAPPs/R/data/Report_Template.docx
    keep_md: yes
    toc: yes
    fig_caption: yes
    fig_width: 12
    fig_height: 6
---

```{r setup_hs67, include=FALSE}

knitr::opts_chunk$set(results = "asis",
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      autodep = TRUE)

options(kableExtra.auto_format = F)

```

```{r, hs67_tbl, echo=FALSE}

hs67 <- readxl::read_xlsx(paste0(data.dir, "hs_tbl.xlsx"), sheet = "hs67")

hs67.tbl <- knitr::kable(hs67, format = "pandoc", padding = 2,
                           caption = tbls(name = "hs67", 
                                          caption = paste0("Summary of the model inputs that are different between Heat Source version 6 and Heat Source version 7.")))

```

### Heat Source version 7

The model parameters for Heat Source version 7 are similar to Heat Source version 6 with a few notable exceptions:

*	Model period can be up to 21 days where Heat Source version 6 only models a single day;       
*	Star pattern landcover input with variable landcover height, density, and ground elevation inputs;       
*	allows for variable flow rate time series on the boundary conditions and tributary inputs;        
*	requires input of latitude, longitude, and aspect for each node of the model;       
*	uses Manning’s equation exclusively to calculate channel hydraulics and omits the ability to specify rating curve coefficients for certain aspects of channel hydraulics;       
*	includes cloudiness (as a percentage of clear sky) as a meteorological input—Heat Source version 6 assumes the clear sky conditions;        
*	allows specifically for groundwater (accretion) and diversion inputs to the model; and        
*	specifies additional morphology data such as bed sediment parameters and channel gradient.        

`r tbls(name = "hs67", display="cite")` summarizes the list of the model parameters that are different between Heat Source version 6 and Heat Source version 7.  

`r hs67.tbl`
