---
output: 
  word_document:
    reference_docx: T:/Temperature_TMDL_Revisions/model_QAPPs/R/data/rmd_template.docx
    keep_md: yes
    toc: yes
    fig_caption: yes
    fig_width: 12
    fig_height: 6
---

```{r setup_hs68, include=FALSE}

knitr::opts_chunk$set(results = "asis",
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      autodep = TRUE)

options(kableExtra.auto_format = F)

```

```{r, hs68_tbl, echo=FALSE}

hs68 <- readxl::read_xlsx(paste0(data.dir, "hs_tbl.xlsx"), sheet = "hs68")

hs68.tbl <- knitr::kable(hs68, format = "pandoc", padding = 2,
                             caption = tbls(name = "hs68", 
                                            caption = paste0("Summary of the model inputs that are different between Heat Source version 6 and Heat Source version 8.")))

```

### Heat Source version 8

The model parameters for Heat Source version 8 are similar to Heat Source version 6 with a few notable exceptions:

*	Model can simulate an unlimited number of days;       
*	Star pattern landcover input with variable landcover height, density, and ground elevation inputs;        
*	allows for variable flow rate time series on the boundary conditions and tributary inputs;        
*	requires input of latitude, longitude and aspect for each node of the model;        
*	uses Manning’s equation exclusively to calculate channel hydraulics and omits the ability to specify rating curve coefficients for certain aspects of channel hydraulics;       
*	includes cloudiness (as a percentage of clear sky) as a meteorological input—Heat Source version 6 assumes the clear sky conditions;        
*	allows specifically for groundwater (accretion) and diversion inputs to the model;        
*	specifies additional morphology data such as bottom width, bed sediment parameters and channel gradient;        
*	specifies bed conduction inputs such as hyporheic exchange parameters; and        
*	allows for the use of LiDAR data to be used for vegetation density and overhang.        

`r tbls(name = "hs68", display="cite")` summarizes the list of model parameters that are different between Heat Source version 6 and Heat Source version 8.

`r hs68.tbl`