---
output: 
  word_document:
    reference_docx: T:/Temperature_TMDL_Revisions/model_QAPPs/R/data/rmd_template.docx
    keep_md: yes
    toc: yes
    fig_caption: yes
    fig_width: 12
    fig_height: 6
---

```{r setup_hs78, include=FALSE}

knitr::opts_chunk$set(results = "asis",
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      autodep = TRUE)

options(kableExtra.auto_format = F)

```

```{r, hs78_tbl, echo=FALSE}

hs78 <- readxl::read_xlsx(paste0(data.dir, "hs_tbl.xlsx"), sheet = "hs78")

hs78.tbl <- knitr::kable(hs78, format = "pandoc", padding = 2,
                             caption = tbls(name = "hs78", 
                                            caption = paste0("Summary of the model inputs and parameters that are different between Heat Source version 7 and Heat Source version 8.")))

```

### Heat Source version 8

The model inputs and parameters for Heat Source version 8 are similar to Heat Source version 7 with a few notable exceptions:        

•	can simulate an unlimited number of days where Heat Source version 7 can model up to 21 days;       
•	specifies different bed sediment parameters such as thermal conductivity, thermal diffusivity, and hyporheic exchange; and        
•	Heat Source version 8 allows for the use of LiDAR data to be used for vegetation density and overhang.        

`r tbls(name = "hs78", display="cite")` summarizes the list of model parameters that are different between Heat Source version 7 and Heat Source version 8. Columns V7 and V8 indicate if the parameter or input is included in version 7 or version 8 of Heat Source. The calibration parameter column indicates if the input or parameter might be modified to improve the calibration.

`r hs78.tbl`
