---
output: 
  word_document:
    reference_docx: T:/Temperature_TMDL_Revisions/model_QAPPs/R/data/Report_Template.docx
    keep_md: yes
    toc: yes
    fig_caption: yes
    fig_width: 12
    fig_height: 6
---

```{r setup_hs8, include=FALSE}

knitr::opts_chunk$set(results = "asis",
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      autodep = TRUE)

options(kableExtra.auto_format = F)

```

```{r, hs8_tbl, echo=FALSE}

hs8 <- readxl::read_xlsx(paste0(data.dir, "hs_tbl.xlsx"), sheet = "hs8")

hs8.tbl <- knitr::kable(hs8, format = "pandoc", padding = 2,
                        caption = tbls(name = "hs8", 
                                       caption = paste0("Summary of model inputs required for Heat Source version 8.")))

```

### Heat Source version 8

`r tbls(name = "hs8", display="cite")` summarizes all of the user entered model inputs and input categories required to run Heat Source version 8; and identifies the subset of inputs that could possibly be modified to improve the calibration of the model. It should be noted, it is unlikely all of these will be used as calibration parameters; rather this list identifies the candidate model inputs that will be considered for adjustment through the calibration process.
The following bulleted list of input categories and specific inputs describes the general form and function of the inputs, and why the inputs are candidates for adjustment during calibration:

*	Morphology – The morphology inputs that could be used as calibration parameters fall into two categories: channel hydraulics and bed conduction. The reasons for inclusion as calibration parameters are described as follows:    

    +	Channel hydraulics – These inputs include stream gradient, bottom width, side slope angle, and Manning’s *n*. Channel hydraulics are important for predicting stream temperatures because they govern the surface area of water that could be exposed to solar radiation, the residence time for exposure, and the degree of light penetration into the water column. Field data for these inputs are often difficult to collect over large spatial scales, and values can vary significantly on a small scale. Heat Source is a one-dimensional model and complex channel configurations are represented as a trapezoidal pattern. Adjustments to inputs that affect channel hydraulic are often necessary to calibrate the model.        
  
    +	Bed conduction – These inputs include hyporheic zone thickness, percent hyporheic exchange, and porosity. Bottom width and side slope angle also affect these inputs by controlling the wetted perimeter of the channel (i.e., the portion or lateral length of the channel bed in direct contact with the stream). These stream morphological characteristics largely govern heat and mass transfer across the stream bed. Typically information on the waterbody sediment size class (e.g. bedrock, gravel, sand, silt) is used as the basis for selecting literature values for these inputs.    

*	Meteorology – The two meteorological inputs typically modified in calibration are percent cloudiness and wind speed. Both cloudiness and wind speed can vary significantly on a small geographic scale and the distance to the source of the meteorological data is often much greater than the small-scale localized weather. Hence, adjusting wind and cloudiness is an appropriate calibration method to account for more site-specific weather patterns.    

*	Mass and thermal flux – Mass and thermal inflows and outflows are inputs often adjusted during the calibration process. These inflows of heat and water consist of tributary and groundwater inflows as well as diversions (i.e., water rights withdrawals) and groundwater losses. The temporal and geographic extents of flow gaging and temperature monitoring on tributaries or groundwater are generally sparse. An effective way of improving the calibration is to complete a flow mass balance with from available data, and then add, subtract, or adjust flows either globally or in specific locations within the bounds of the flow mass balance and available measurements, and the temperature response predicted by the model.    

*	Vegetation – Vegetation characteristics input into the model are often derived from aerial imagery or LiDAR. The vegetation characteristics determine the degree to which near-stream vegetation has the capacity to block incidental solar radiation on the surface of the modeled waterbody. Three vegetation inputs incorporated into the model calibration process are the vegetation density, overhang, and height. Field measurements offer a general understanding of vegetation characteristics within the watershed, however variability in these parameters can be significant on smaller geographic scales. To improve the model fit these model inputs may be modified on a global scale for different vegetation classes within the bounds of available data.

`r hs8.tbl`