---
output: 
  word_document:
    reference_docx: T:/Temperature_TMDL_Revisions/model_QAPPs/R/data/rmd_template.docx
    keep_md: yes
    toc: yes
    fig_caption: yes
    fig_width: 12
    fig_height: 6
---

```{r setup_hs9, include=FALSE}

knitr::opts_chunk$set(results = "asis",
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      autodep = TRUE)

options(kableExtra.auto_format = F)

```

```{r hs9_tbl, echo=FALSE}

hs9.v <- tmdl.mod.2 %>% 
  dplyr::filter(mod_rmd == "hs9") %>% 
  pull(Model_version)

hs9 <- readxl::read_xlsx(paste0(data.dir, "hs_tbl.xlsx"), sheet = "hs9")

hs9.tbl <- knitr::kable(hs9, format = "pandoc", padding = 2,
                        caption = tbls(name = "hs9", 
                                       caption = paste0("Summary of model inputs required for Heat Source version 9.")))

```

### `r unique(hs9.v)`

The Heat Source version 9 shade module is a sub program within Heat Source version 9 used to model effective shade and solar radiation flux. Heat Source 9 is similar to other versions of heat source, especially for the landcover and morphology related inputs used by the shade module. `r tbls(name = "hs9", display="cite")` summarizes all of the user entered model inputs required to run the shade module of Heat Source version 9. `r tbls(name = "hs9", display="cite")` also identifies the subset of inputs that could possibly be modified to improve the calibration of the model. It should be noted, it is unlikely all of these will be used as calibration parameters; rather this list identifies the candidate model inputs that will be considered for adjustment through the calibration process.

`r hs9.tbl`