---
author: "Ryan Michie"
date: "`r format(Sys.Date(), '%m/%d/%Y')`"
output:
  word_document:
    reference_docx: N:/Status_and_Trend_Reports/Report_Files/Report_Template.docx
    keep_md: yes
    toc: yes
    fig_caption: yes
    fig_width: 12
    fig_height: 6
  html_document:
    toc: yes
    df_print: paged
---

---
title: "`r paste0("Modeling Quality Assurance Project Plan for the ", qapp_project_area, " Temperature Total Maximum Daily Load")`"
---

```{r, label=`setup`, include=FALSE}
options(knitr.duplicate.label = 'allow')
options(knitr.kable.NA = '')
options(tinytex.verbose = TRUE)

knitr::opts_chunk$set(results = "asis",
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      autodep = TRUE)

library(knitr); library(kableExtra)
library(tidyverse)
library(captioner)
library(readxl)
library(lubridate)
library(flextable)
library(sf); library(sp)
library(rgdal)
library(ggplot2)
library(stringr)
library(english)
library(pacman)

tbls  <- captioner::captioner(prefix="Table")
figs <- captioner::captioner(prefix="Figure")
eqns <- captioner::captioner(prefix="Equation")

```

# Introduction

This Quality Assurance Project Plan (QAPP) is a prescriptive outline for conducting analysis and modeling to address temperature impairments within the `r qapp_project_area`. The `r qapp_project_area` includes several waterbodies listed on the Oregon 2018/2020 Section 303(d) Category 5 list as water quality limited for temperature. To address these impairments, the Oregon Department of Environmental Quality (DEQ) will develop a Total Maximum Daily Load (TMDL) to address anthropogenic sources or causes of pollution contributing to water quality impairments. For the `r qapp_project_area`, a temperature TMDL was previously developed and implemented, and this TMDL will constitute a revision of that TMDL. 

To aid in this process, this QAPP has been developed as a guide to conducting the analysis and numerical modeling that will support the TMDL. To those ends, the modeling and analysis used in the previous TMDL will form the basis of the technical work for this TMDL revision. This QAPP has been developed to inform the project team what information to implement and how from the previous TMDL, as well as how to incorporate new information since the issuance of the original TMDL. In particular, this QAPP details the following:

*	Definition of the issue and objectives, including the spatial and temporal extents of the water quality impairments (Section 2);    

*	A high-level assessment of sources of pollution (Section 3);    

*	The overarching technical approach, including the appropriate modeling and analytical tools to be used (Section 4);    

*	The data sources for defining and creating inputs to the model, including data that were used in the modeling for the original TMDL. Examples of these inputs include weather, tributary stream flow and temperature, point sources and vegetation characteristics (Sections 5 and 6);    

*	How the analysis and modeling will be evaluated for acceptability (Sections 7 and 9);    

*	Scenarios for evaluating management strategies for reducing anthropogenic heat loads (Section 10);     

*	Various aspects for managing the TMDL development project, including documentation (Section 8), the project team (Section 11), data and records management (Sections 12  and 13); and    

*	Aspects relating to this QAPP and its role in the project (Sections 14  and 15).    


# Problem Definition and Management Objectives

```{r, label=`cat45`, echo=FALSE}

t2 <- pro.cat.45.rivers.tbl %>% 
  dplyr::group_by(AU_Name,IR_categor,Year_liste) %>% 
  dplyr::summarize(Period = toString(unique(sort(Period)))) %>% 
  dplyr::ungroup() %>% 
  dplyr::rename(Waterbody = AU_Name,
                Category = IR_categor,
                `Year Listed` = Year_liste,
                `Criterion Period` = Period) %>% 
  dplyr::select(Waterbody,`Year Listed`,Category,`Criterion Period`)

if(NROW(t2)>0){
  
  t2 <- knitr::kable(t2, format = "pandoc", padding = 2,
                            caption = tbls(name = "t2",
                                           caption = paste0("Waterbodies listed for temperature impairment based on Section 303(d) in the 2018/20 Integrated Report")))
  
}

```

In 2007 the TMDL for the North Umpqua Subbasin was developed for bacteria, aquatic weeds, biological criteria and temperature. The original temperature TMDLs were based on the Natural Conditions Criteria (NCC) identified in Oregon Administrative Rules (OAR), specifically OAR 340-041-0028(8). These criteria were invalidated by the Oregon U.S. District Court. The applicable numeric and narrative temperature criteria must be used. As required by court order, the U.S. Environmental Protection Agency (USEPA) and State of Oregon (OR) are required to revise the water temperature TMDLs for the North Umpqua Subbasin. In revising the TMDLs the models used for the North Umpqua Subbasin mainstem will be updated to target the biologically-based numeric criteria (BBNC) and Human Use Allowance (HUA) WQS. The tributary segment TMDLs will also be updated with the BBNC and HUA WQS and a thermal load calculation.

The existing temperature TMDL addresses rearing/migration and spawning period temperature impairments on streams without point sources or dams for the entire basin. In rivers that have no point sources or dams, activities designed to improve summer stream temperatures are the same activities that will improve fall and winter temperatures. The nonpoint source load allocations are expressed as effective shade targets and apply year-round.

Furthermore, the existing temperature TMDL addresses streams with point sources and dams during the rearing/migration (i.e. non-spawning) time period. Waste load allocations have been developed for point sources during the portion of the year when spawning does not occur and will be incorporated into the NPDES permits. During the spawning period, there are three impaired segments on the North Umpqua River that are downstream of a hydro-electric project and a point source is present. More data and analysis are needed to complete those TMDLs. Likewise, on other streams and rivers that are not currently identified as impaired during the spawning period and have point sources or dams, the TMDLs were not computed.

Since the issuance of the original TMDL, the waterbody impairments may have changed based on the 2018/2020 Section 303(d) Category 5 list. These changes may affect the location(s) and time(s) of year. As part of the TMDL update, DEQ will address the temperature impairments for these water bodies based on the 2018/2020 list. Those listings, as they pertain to the `r subbasin` QAPP project area, are summarized in `r tbls(name = "t2", display="cite")`. 

`r t2`

This QAPP document provides the management context and information about existing models to be used for the TMDL revisions.

# Conceptual Model: Key Processes and Variables

Anthropogenic heat sources are derived from solar radiation as increased levels of sunlight reach the stream surface and effluent discharges to surface waters. Therefore, the pollutants targeted in the existing TMDL are: 

**Anthropogenic Nonpoint Sources**: Heat from human-caused solar radiation loading increases to the stream network, as a result of alterations in near stream vegetation, channel morphology, and flow modifications.

Historically, human activities have altered the stream morphology and hydrology and decreased the amount of riparian vegetation in the basin. The basin includes urban, agricultural, and forested lands. Additionally, hydroelectric projects and multiple points of diversion in the Umpqua River Basin have altered stream flow levels. 

Low summertime flows decrease the thermal assimilative capacity of streams. Pollutant (solar radiation) loading causes larger temperature increases in stream segments where flows are reduced by human uses. 

Larger streams, such as the North Umpqua Rivers naturally experience a large solar flux because their channels are too wide for riparian vegetation to effectively shade the surface of the water. Smaller tributaries, such as Little River, are easier to shade and have smaller background solar flux values. Additionally, effective shade levels on smaller streams are more sensitive to riparian disturbances and so the differences between current condition solar flux and background solar flux can be larger.

**Anthropogenic Point Sources**: Heat from warm water discharges of human origin, such as industrial outfalls, municipal waste-water treatment plants (WWTP), and other point sources.

PacifiCorp’s hydro electric project on the North Umpqua River is responsible for elevated stream temperatures between Lemolo Reservoir and the Umpqua River tidewater boundary.

# Technical Approach

## Overview

Stream temperature TMDLs are generally scaled to a subbasin or basin and include all perennial surface waters that have salmonid presence or that contribute to areas with salmonid presence. Since stream temperatures are affected by cumulative interactions between upstream and local sources, the TMDL considers all surface waters that affect the temperatures of 303(d) listed waterbodies. For example, the `r subbasin` is water quality limited for temperature. To address this listing in the TMDL, all major tributaries are included in the TMDL analysis and TMDL allocations are applied throughout the entire stream network.

An important step in the TMDL is to perform a source assessment which quantifies the anthropogenic contributions to stream heating. One anthropogenic contribution to solar radiation heat loading results from decreased stream surface shade. Decreased stream shade may be caused by near stream vegetation disturbance/removal and channel morphology changes. Other anthropogenic sources of stream warming may include stream flow reductions and warm water point source effluent.

The pollutant addressed in the temperature TMDL is heat. Anthropogenic nonpoint and point sources are not permitted to heat a waterbody more than 0.3 $^\circ C$ above the applicable criteria, cumulatively at the point of maximum impact. Allocated conditions are expressed as solar heat load and solar heat flux (watts, and watts per square meter, respectively). Nonpoint source heat allocations are translated into effective shade surrogate measures. Effective shade surrogate measures provide site-specific targets for land managers. Attainment of the surrogate measures ensures compliance with the nonpoint source allocations. Point source waste load allocations are based on the applicable numeric and/or narrative criteria. Point sources are not allowed to increase stream temperatures more than 0.1 $^\circ C$ (a portion of the 0.3 $^\circ C$ human use allowance) cumulatively at the point of maximum impact.

Stream temperatures were simulated using a computer model (Heat Source) for the main rivers and their larger tributaries. Simulations focus on the larger streams that contain or influence primary fish habitat. Site-specific load allocations have been developed for the streams that were simulated. Other streams are assigned generalized load allocations based on potential vegetation and effective shade.

## Model Selection

The modeling framework needs for this project include:

1)	Prediction of hourly stream temperatures over a period of months and at a no greater than 500 meter longitudinal resolution.  

2)	Prediction of hourly solar radiation flux and daily effective shade at a no greater than 50 meter longitudinal resolution.  

3)	Ability to evaluate hourly stream temperature response from changes in streamside vegetation within the upstream catchment.  

4)	Ability to evaluate hourly stream temperature response from changes in water withdrawals and tributary stream flow within the upstream catchment.  

5)	Ability to evaluate hourly stream temperature response from changes in channel morphology within the upstream catchment.  

6)	Ability to evaluate hourly stream temperature response from changes in effluent temperature and flow discharge from NPDES permitted facilities.  

7)	Prediction of hourly water temperatures in stratified reservoir systems along the North Umpqua River system over a period of months.  

8)	Ability to evaluate hourly stream and reservoir temperature response from changes in dam operations along the North Umpqua River System.  

For these reasons, the Heat Source stream thermodynamics model was selected for stream temperature simulation.  

# Data Availability and Quality

## Meteorology

```{r, label=`t51`, echo=FALSE}

if(NROW(model.input[which(model.input$`Model Location Type` == "Meteorological"),])>0){
  
  t51_sp <- model.input %>% 
    dplyr::filter(`Model Location Type` %in% c("Meteorological")) %>%
    dplyr::filter(!is.na(`Data Source`) & is.na(`Interpolated Data`)) %>%
    dplyr::mutate(Station = ifelse(`Station ID` == "No Station ID or unknown" | is.na(`Station ID`),
                                   `Data Source`,
                                   paste0(`Station ID`, ", ", `Data Source`)),
                  `Latitude/Longitude` = ifelse(is.na(Latitude), NA, paste0(round(Latitude,4), ", ",round(Longitude,3))),
                  `Available Data` = Parameter) %>% 
    dplyr::select(Station, `Latitude/Longitude`, `Available Data`) %>%
    dplyr::group_by(Station, `Latitude/Longitude`) %>% 
    dplyr::summarize(`Available Data` = toString(unique(sort(`Available Data`)))) %>% 
    dplyr::ungroup()
  
  if(NROW(t51_sp)>0){
    
    t51_sp_tbl <- knitr::kable(t51_sp, format = "pandoc", padding = 2,
                               caption = tbls(name = "t51_sp",
                                              caption = paste0("Meteorological stations and data available in DEQ files for the ", qapp_project_area, ".")))
    
  }
  
}

t51_ncdc <- ncdc.station.tbl %>% 
  dplyr::filter(!str_detect(station.id,"COOP")) %>% 
  dplyr::mutate(`Station` = name,
                `Station ID` = station.id,
                `Latitude/Longitude` = paste0(round(lat,4), ", ",round(long,3)),
                `Available Data` = Parameter) %>%
  dplyr::group_by(`Station`, `Station ID`, `Latitude/Longitude`, `Available Data`) %>% 
  dplyr::summarize(n=n()) %>%
  dplyr::ungroup() %>% 
  dplyr::select(`Station`, `Station ID`, `Latitude/Longitude`, `Available Data`) %>% 
  dplyr::arrange(`Station`, `Station ID`)

if(NROW(t51_ncdc)>0){
  
  t51_ncdc_tbl <- knitr::kable(t51_ncdc, format = "pandoc", padding = 2,
                               caption = tbls(name = "t51_ncdc",
                                              caption = paste0("Meteorological stations and data available in the National Climatic Data Center (NCDC) database in the ",
                                                               qapp_project_area, ".")))
}

t51_raws <- raws.station.tbl %>% 
  dplyr::mutate(`Station` = siteName,
                `Station wrccID` = wrccID,
                `Latitude/Longitude` = paste0(round(lat,4), ", ",round(long,3)),
                Agency = agency) %>% 
  dplyr::group_by(`Station`, `Station wrccID`, `Latitude/Longitude`, Agency) %>% 
  dplyr::summarize(n=n()) %>%
  dplyr::ungroup() %>%
  dplyr::select(`Station`, `Station wrccID`, `Latitude/Longitude`, Agency) %>% 
  dplyr::arrange(`Station`, `Station wrccID`)

if(NROW(t51_raws)>0){
  
  t51_raws_tbl <- knitr::kable(t51_raws, format = "pandoc", padding = 2,
                               caption = tbls(name = "t51_raws",
                                              caption = paste0("Meteorological stations and data, including ", knitr::combine_words(unique(sort(tolower(raws.station.tbl$Parameter)))), ", available in the Remote Automatic Weather Station (RAWS) database in the ", qapp_project_area, ".")))
  
}

t51_agrimet <- agrimet.station.tbl %>%
  dplyr::mutate(`Station` = description.x,
                `Station ID` = siteid,
                `Latitude/Longitude` = paste0(round(lat,4), ", ",round(long,3))) %>% 
  dplyr::group_by(`Station`, `Station ID`, `Latitude/Longitude`) %>% 
  dplyr::summarize(n=n()) %>%
  dplyr::ungroup() %>%
  dplyr::select(`Station`, `Station ID`, `Latitude/Longitude`) %>% 
  dplyr::arrange(`Station`, `Station ID`)

if(NROW(t51_agrimet)>0){
  
  t51_agrimet_tbl <- knitr::kable(t51_agrimet, format = "pandoc", padding = 2,
                                  caption = tbls(name = "t51_agrimet",
                                                 caption = paste0("Meteorological stations and data, including air temperature, precipitation, relative humidity and wind, available in the USBL AgriMet database in the ", qapp_project_area, ".")))
  
}

t51_hydromet <- hydromet.station.tbl %>% 
  dplyr::filter(Parameter %in% c("Air Temperature","Precipitation")) %>% 
  dplyr::mutate(`Station` = Station.Name,
                `Station ID` = Station.ID,
                `Latitude/Longitude` = paste0(round(as.numeric(Lat),4), ", ",round(as.numeric(Long),3)),
                `Available Data` = Parameter) %>% 
  dplyr::select(`Station`, `Station ID`, `Latitude/Longitude`, `Available Data`) %>%
  dplyr::group_by(`Station`, `Station ID`, `Latitude/Longitude`) %>% 
  dplyr::summarize(`Available Data` = toString(unique(sort(`Available Data`)))) %>% 
  dplyr::ungroup()

if(NROW(t51_hydromet)>0){
  
  t51_hydromet_tbl <- knitr::kable(t51_hydromet, format = "pandoc", padding = 2,
                                   caption = tbls(name = "t51_hydromet",
                                                  caption = paste0("Meteorological stations and data available in the USBL Hydromet database in the ", qapp_project_area, ".")))
  
}

t51_mw <- mw.station.tbl %>% 
  dplyr::mutate(`Station` = NAME,
                `Station ID` = STID,
                `Latitude/Longitude` = paste0(round(as.numeric(lat),4), ", ",round(as.numeric(long),3))) %>% 
  dplyr::group_by(`Station`, `Station ID`, `Latitude/Longitude`) %>% 
  dplyr::summarize(n=n()) %>%
  dplyr::ungroup() %>%
  dplyr::select(`Station`, `Station ID`, `Latitude/Longitude`) %>% 
  dplyr::arrange(`Station`, `Station ID`)

if(NROW(t51_mw)>0){
  
  t51_mw_tbl <- knitr::kable(t51_mw, format = "pandoc", padding = 2,
                             caption = tbls(name = "t51_mw",
                                            caption = paste0("Meteorological stations and data, including air temperature, precipitation, relative humidity, wind speed and wind direction, available in the MesoWest database in the ", qapp_project_area, ".")))
  
}

if(NROW(model.input[which(model.input$`Model Location Type` == "Meteorological"),])>0){
  
  tables.met <- data.frame(tbl_name = c("t51_sp","t51_ncdc","t51_raws","t51_agrimet","t51_hydromet","t51_mw"),
                           tbl_rows = c(NROW(t51_sp),NROW(t51_ncdc),NROW(t51_raws),NROW(t51_agrimet),NROW(t51_hydromet),NROW(t51_mw))) %>% 
    dplyr::filter(tbl_rows != "0")
  
  t51_num <- NROW(tables.met)
  
} else {
  
  tables.met <- data.frame(tbl_name = c("t51_ncdc","t51_raws","t51_agrimet","t51_hydromet","t51_mw"),
                           tbl_rows = c(NROW(t51_ncdc),NROW(t51_raws),NROW(t51_agrimet),NROW(t51_hydromet),NROW(t51_mw))) %>% 
    dplyr::filter(tbl_rows != "0")
  
  t51_num <- NROW(tables.met)
  
}

```

Meteorological data, including air temperature, cloudiness, relative humidity, and wind speed, are available from ODEQ database, NOAA National Climatic Data Center (NCDC), National Interagency Fire Center (NIFC) Remote Automatic Weather Stations (RAWS), Bureau of Reclamation Cooperative Agricultural Weather Network (AgriMet) and Automated Hydrologic and Meteorologic Monitoring Stations Network (Hydromet), and University of Utah MesoWest database. `r ifelse(NROW(model.input[which(model.input$"Model Location Type" == "Meteorological"),])>0,tbls(name = "t51_sp", display="cite"),tbls(name = "t51_ncdc", display="cite"))` - Table `r t51_num` list available meteorological data from these sources in the `r qapp_project_area`.

`r if(NROW(model.input[which(model.input$"Model Location Type" == "Meteorological"),])>0){t51_sp_tbl}`
`r if(NROW(t51_ncdc)>0){t51_ncdc_tbl}`
`r if(NROW(t51_raws)>0){t51_raws_tbl}`
`r if(NROW(t51_agrimet)>0){t51_agrimet_tbl}`
`r if(NROW(t51_hydromet)>0){t51_hydromet_tbl}`
`r if(NROW(t51_mw)>0){t51_mw_tbl}`

## Thermal Infrared (TIR) Data

```{r, label=`t52`, echo=FALSE}

t52 <- model.input %>% 
  dplyr::filter(`Station ID` == "TIR") %>% 
  dplyr::rename(`Abbreviated Reference` = `Data Source`) %>% 
  dplyr::left_join(ref, by = "Abbreviated Reference") %>% 
  dplyr::mutate(`Abbreviated Reference` = strip_alpha(`Abbreviated Reference`)) %>% 
  dplyr::mutate(Reference = paste(`Full Reference`, ifelse(is.na(`Reference Link`),"",`Reference Link`)))

waterbodies <- unique(sort(t52$`Model Waterbody`))

citations <- unique(sort(t52$`Abbreviated Reference`))

```

ODEQ contracted with Watershed Sciences, Inc. to provide airborne thermal infrared (TIR) imagery of spatial temperature patterns within the `r qapp_project_area`. TIR data were used to characterize the thermal regime of the streams and habitat quality in the basin where the TIR data were available.

`r paste0("TIR data are available for ", knitr::combine_words(waterbodies), " (", knitr::combine_words(citations), ").")`

## Continuous Stream Temperature Data

```{r, label=`t53`, echo=FALSE}

st_awqms <- station_awqms %>% 
  dplyr::rename(`Station Description` = StationDes,
                Organization = OrgID,
                Latitude = Lat_DD,
                Longitude = Long_DD) %>% 
  dplyr::select(`Station Description`,`Station ID`, Organization, Latitude, Longitude)

st_model <- station_model %>% 
  dplyr::filter(`Parameter` %in%  c("Water Temperature")) %>% 
  dplyr::filter(!is.na(`Data Source`) & is.na(`Interpolated Data`)) %>% 
  dplyr::mutate(`Station Description` = ifelse(is.na(StationDes),`Model Location Name`,StationDes),
                Organization = ifelse(is.na(OrgID),`Data Source`,OrgID)) %>% 
  dplyr::select(`Station Description`,`Station ID`, Organization, Latitude, Longitude)

t53_1 <- rbind(st_awqms,st_model) %>% 
  dplyr::filter(!`Station ID` == "TIR") %>% 
  dplyr::mutate(`Station Name and ID` = ifelse(`Station ID` == "No Station ID or unknown" | is.na(`Station ID`),
                                               `Station Description`,
                                               paste0(`Station Description`, " (", `Station ID`, ")"))) %>% 
  dplyr::mutate(`Latitude/Longitude` = ifelse(is.na(Latitude), NA, paste0(round(Latitude,4), ", ",round(Longitude,3)))) %>% 
  dplyr::mutate(`Station Name and ID` = stringr::str_to_title(`Station Name and ID`)) %>% 
  dplyr::mutate_at("Station Name and ID", str_replace_all, "Ordeq", "ORDEQ") %>%
  dplyr::mutate_at("Station Name and ID", str_replace_all, " Rm", " RM") %>%
  dplyr::mutate_at("Station Name and ID", str_replace_all, "Lb", "LB") %>%
  dplyr::mutate_at("Station Name and ID", str_replace_all, "Nf", "NF") %>%
  dplyr::mutate_at("Station Name and ID", str_replace_all, "Sf", "SF") %>%
  dplyr::distinct(`Station Name and ID`, .keep_all=TRUE) %>%
  dplyr::mutate(Organization = ifelse(Organization == "OregonDEQ", "ODEQ", Organization)) %>% 
  dplyr::mutate(Organization = ifelse(Organization == "11NPSWRD_WQX", "EPA WQX", Organization)) %>% 
  dplyr::mutate(Organization = ifelse(Organization == "USFS(NOSTORETID)", "USFS", Organization)) %>%
  dplyr::select(`Station Name and ID`,`Latitude/Longitude`, Organization) %>% 
  dplyr::arrange(`Station Name and ID`)

if(NROW(t53_1)>0){
  
  t53_1_tbl <-  knitr::kable(t53_1, format = "pandoc", padding = 2,
                             caption = tbls(name = "t53_1",
                                            caption = paste0("Continuous temperature monitoring stations in the ", qapp_project_area, " currently available in public databases and DEQ files.")))
  
}

t53_2 <- data.sample.count %>%
  dplyr::filter(!`Station ID` == "TIR") %>% 
  dplyr::mutate(`Station Name and ID` = ifelse(`Station ID` == "No Station ID or unknown" | is.na(`Station ID`),
                                               StationDes,
                                               paste0(StationDes, " (", `Station ID`, ")"))) %>% 
  dplyr::select(Year, `Station Name and ID`, Jan, Feb, Mar, Apr, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec) %>% 
  dplyr::arrange(Year, `Station Name and ID`)

if(NROW(t53_2)>0){
  
  t53_2_tbl <- knitr::kable(t53_2, format = "pandoc", padding = 2,
                            caption = tbls(name = "t53_2",
                                           caption = paste0("Summary of existing temperature data in the ", qapp_project_area,". Columns Jan – Dec indicate the number of daily maximum temperature results in each month.")))
  
}


```

All available continuous stream temperature data were retrieved from DEQ’s Ambient Water Quality Monitoring System (AWQMS). `r tbls(name = "t53_1", display="cite")` list all the locations where continuous stream temperature data were collected and the organizations that collected that data. `r tbls(name = "t53_2", display="cite")` summarizes when data were collected at each station.

`r t53_1_tbl`
`r t53_2_tbl`

## Stream Flow Data

```{r, label=`t54`, echo=FALSE}

t54_usgs <- usgs.station.tbl %>% 
  dplyr::mutate(Station = station_nm,
                `Station ID`= site_no,
                `Latitude/Longitude` = paste0(round(dec_lat_va,4), ", ",round(dec_long_va,3))) %>% 
  dplyr::select(Station, `Station ID`, `Latitude/Longitude`) %>% 
  dplyr::arrange(Station)

t54_hydromet <- hydromet.station.tbl %>% 
  dplyr::filter(Parameter %in% c("Discharge", "Spillway Discharge", "Flow")) %>% 
  dplyr::distinct(Lat, Long, .keep_all = TRUE) %>% 
  dplyr::mutate(Station = Station.Name,
                `Station ID` = Station.ID,
                `Latitude/Longitude` = paste0(round(as.numeric(Lat),4), ", ",round(as.numeric(Long),3))) %>% 
  dplyr::select(Station, `Station ID`, `Latitude/Longitude`) %>% 
  dplyr::arrange(Station)

t54_sp <- model.input %>% 
  dplyr::filter(`Parameter` %in% c("Flow")) %>%
  dplyr::filter(!`Model Location Type` == "Point of Diversion") %>% 
  dplyr::filter(!`Data Source` == "USGS" & is.na(`Interpolated Data`)) %>% 
  dplyr::distinct(Latitude, Longitude, .keep_all = TRUE) %>% 
  dplyr::mutate(Station = `Model Location Name`,
                `Latitude/Longitude` = ifelse(is.na(Latitude), NA, paste0(Latitude, ", ", Longitude))) %>% 
  dplyr::select(Station, `Station ID`, `Latitude/Longitude`, `Data Source`) %>% 
  dplyr::arrange(Station)

if(NROW(t54_usgs)>0){
  
  t54_usgs_tbl <- knitr::kable(t54_usgs, format = "pandoc", padding = 2,
                               caption = tbls(name = "t54_usgs",
                                              caption = paste0("Continuous flow measurements available from the USGS flow gaging stations in the ", qapp_project_area, ".")))
  
}

if(NROW(t54_hydromet)>0){
  
  t54_hydromet_tbl <- knitr::kable(t54_hydromet, format = "pandoc", padding = 2,
                                   caption = tbls(name = "t54_hydromet",
                                                  caption = paste0("Continuous flow measurements available from the USBR Hydromet monitoring stations in the ", qapp_project_area, ".")))
  
}

if(NROW(t54_sp)>0){
  
  t54_sp_tbl <- knitr::kable(t54_sp, format = "pandoc", padding = 2,
                             caption = tbls(name = "t54_sp",
                                            caption = paste0("Flow measurements from ODEQ and other public databases in the ", qapp_project_area, ".")))
  
  t54.sp.tbl.num <- tbls(name = "t54_sp", display="cite")
  
}

```

`r tbls(name = "t54_usgs", display="cite")` - `r tbls(name = "t54_sp", display="cite")` list the stations for flow volume data and other instream measurements that were available in USGS, Oregon DEQ, and other public databases.

`r t54_usgs_tbl`
`r t54_hydromet_tbl`
`r t54_sp_tbl`

## Point Source Discharges

```{r, label=`t55`, echo=FALSE}

section.npdes.ind <- npdes.ind %>% 
  dplyr::filter(`Project Area` %in%  qapp_project_area) %>%
  dplyr::mutate(`Common Name` = stringr::str_to_title(`Common Name`)) %>%
  dplyr::mutate_at("Common Name", str_replace_all, "Stp", "STP") %>%
  dplyr::mutate(`Facility Name and Number` = paste0(`Common Name`," (", `WQ File Nbr`,")"),
                `Latitude/Longitude` = paste0(round(Latitude,4), ", ",round(Longitude,3)),
                `Permit Type and Description` = paste0(`Permit Type`, ": ", `Permit Description`),
                `Stream/River Mile` = ifelse(is.na(`Stream Name`), NA, paste0(`Stream Name`, " ", " RM ",`River Mile`))) %>% 
  dplyr::select(`Facility Name and Number`, `Latitude/Longitude`,`Permit Type and Description`, `Stream/River Mile`) %>% 
  dplyr::arrange(`Facility Name and Number`)

if(NROW(section.npdes.ind)>0){
  
  t55_ind_tbl <- knitr::kable(section.npdes.ind, format = "pandoc", padding = 2,
                              caption = tbls(name = "t55_ind",
                                             caption = paste0("Summary of individual NPDES permitted discharges on ", qapp_project_area, ".")))
  
}

section.npdes.gen <- npdes.gen %>% 
  dplyr::filter(`Project Area` %in%  qapp_project_area)%>%
  dplyr::mutate(`Permit Type and Description` = paste0(PermitType, ": ", PermitDescription)) %>% 
  dplyr::group_by(`Permit Type and Description`) %>% 
  dplyr::summarise(`Number of Permits` = n()) %>% 
  dplyr::ungroup()

if(NROW(section.npdes.gen)>0){
  
  t55_gen_tbl <- knitr::kable(section.npdes.gen, format = "pandoc", padding = 2,
                              caption = tbls(name = "t55_gen",
                                             caption = paste0("Summary of general NPDES permitting types in the ", qapp_project_area, ".")))
  
}

```

`r tbls(name = "t55_ind", display="cite")` identifies all the active individual NPDES permittees in the `r qapp_project_area`. Many of these permittees submit Discharge Monitoring Reports (DMRs) as a condition of their permit. Depending on the monitoring requirements in the permit, some permittees are required to report effluent temperature and effluent flow rates in the DMR. The frequency and type of reporting varies by permit and permit type. Some permits only require monthly, weekly, or daily grab samples while others require summary statistics such as daily maximum, daily mean, or seven day average daily maximum. When possible, DEQ will utilize any continuous effluent data that has been provided to DEQ. When continuous data is not available, DMR data will be utilized to characterize point source discharges. `r tbls(name = "t55_gen", display="cite")` lists all the active general NPDES permitting types in the `r qapp_project_area`.

`r t55_ind_tbl`
`r if(NROW(section.npdes.gen)>0){t55_gen_tbl}`

## Water Rights/Surface Water Diversions

Data on surface water diversion rates (usage) and the points of diversion (location) are available from the Oregon Water Resources Department (OWRD). OWRD regulates all commercial, industrial, domestic, and agricultural water use in the state of Oregon through water rights.

Estimates of water diversion rates and location of points of diversion can be derived from the following OWRD sources:

* [Water Rights Information System][WRIS] (WRIS) – the WRIS database contains all permitted or certificated water rights. Data in the WRIS corresponding to quantities of water for use are expressed as maximum use allowable, generally as monthly, seasonal or annual rates or volumes. These maximum values may not correspond to actual usage, which will likely vary based on factors such as irrigation application rate or household consumer demand. Oregon DEQ may choose to incorporate the maximum amount allowable or some lesser quantity provided sufficient information is available to support those rates in the modeling.  Water rights information can also be accessed using their online mapping application (https://apps.wrd.state.or.us/apps/gis/wr/Default.aspx).
    
* [Water Use Reports][Water Use Reports] – some, but not all, water rights holders must monitor and report the water they use to the state, typically on a monthly or yearly basis, as a requirement of their water rights. These water use reports will be used to develop withdrawal time series based on the information at hand.

[WRIS]: https://www.oregon.gov/owrd/programs/WaterRights/WRIS/Pages/default.aspx
[Water Use Reports]: https://apps.wrd.state.or.us/apps/wr/wateruse_query/

## Effective Shade Measurements

```{r, label=`t56`, echo=FALSE}

t56 <- model.input %>% 
  dplyr::filter(`Parameter` %in% c("Effective Shade")) %>% 
  dplyr::filter(is.na(`Interpolated Data`)) %>% 
  dplyr::distinct(Latitude, Longitude, .keep_all = TRUE) %>% 
  dplyr::mutate(Station = `Model Location Name`,
                `Latitude/Longitude` = ifelse(is.na(Latitude), NA, paste0(Latitude, ", ", Longitude))) %>% 
  dplyr::select(Station, `Station ID`, `Latitude/Longitude`, `Data Source`)

if(NROW(t56)>0){
  
  t56_tbl <- knitr::kable(t56, format = "pandoc", padding = 2,
                          caption = tbls(name = "t56",
                                         caption = paste0("Effective shade data collected in the  ", qapp_project_area, ".")))
  
}

```

`r tbls(name = "t56", display="cite")` lists the stations for effective shade measurements that were available in the public and Oregon DEQ databases.

`r t56_tbl`

# Model Development and Calibration

## General Model Parameters

```{r, label=`chap6`, include=FALSE}

pro.area.tmdls <- knitr::combine_words(unique(model.info$"TMDL Document"))

# summary table
tmdl.mod.1<- model.info %>% 
  dplyr::group_by(Model_version) %>% 
  dplyr::summarize(`Model Waterbody` = toString(unique(sort(`Model Waterbody`)))) %>% 
  dplyr::ungroup() %>% 
  dplyr::rename(`Model Version` = Model_version)

tmdl.mod.tbl <- knitr::kable(tmdl.mod.1, format = "pandoc", padding = 2,
                          caption = tbls(name = "tmdl.mod",
                                         caption = paste0("Models used in the development of the ", pro.area.tmdls, ".")))

tmdl.mod.2 <- model.info %>% 
  dplyr::distinct(Model_version,`QAPP Project Area`,`Primary Model Parameter`,Model_version,mod_rmd,mod_score)

mod.check <- sum(as.numeric(tmdl.mod.2$mod_score))

# model version tables
hs6 <- readxl::read_xlsx(paste0(data.dir, "hs_tbl.xlsx"), sheet = "hs6")  
hs67 <- readxl::read_xlsx(paste0(data.dir, "hs_tbl.xlsx"), sheet = "hs67")
hs68 <- readxl::read_xlsx(paste0(data.dir, "hs_tbl.xlsx"), sheet = "hs68")

if(mod.check == "11"){

  hs67.tbl <- knitr::kable(hs67, format = "pandoc", padding = 2,
                           caption = tbls(name = "hs67", 
                                          caption = paste0("Summary of model inputs and input categories required for Heat Source version 7.")))
  
} else {
  
  if(mod.check == "21"){
    
    hs68.tbl <- knitr::kable(hs68, format = "pandoc", padding = 2,
                             caption = tbls(name = "hs68", 
                                            caption = paste0("Summary of model inputs and input categories required for Heat Source version 8.")))
    
  } else {
    
    if(mod.check == "31"){
      
      hs67.tbl <- knitr::kable(hs67, format = "pandoc", padding = 2,
                               caption = tbls(name = "hs67", 
                                              caption = paste0("Summary of model inputs and input categories required for Heat Source version 7.")))
     
      hs68.tbl <- knitr::kable(hs68, format = "pandoc", padding = 2,
                               caption = tbls(name = "hs68", 
                                              caption = paste0("Summary of model inputs and input categories required for Heat Source version 8.")))  
      
    } else {
      
      hs6.tbl <- knitr::kable(hs6, format = "pandoc", padding = 2,
                              caption = tbls(name = "hs6", 
                                             caption = paste0("Summary of model inputs and input categories required for Heat Source version 6.")))
    }
  }
}

# model versions


mod.v <- NULL

if(mod.check == "11"){
  
  modv <- tmdl.mod.2 %>% 
    dplyr::filter(!mod_rmd %in% c("hs6","hs7"))
  
  for(mod in unique(sort(modv$mod_rmd))){
    
    mod.child <- paste0(mod,".Rmd")

    mod.v <- c(knitr::knit_child(input = "hs6.Rmd", envir = globalenv()),
               knitr::knit_child(input = "hs67.Rmd", envir = globalenv()),
               knitr::knit_child(input = mod.child, envir = globalenv()))
    
  }
  
} else {
  
  if(mod.check == "21"){
    
    modv <- tmdl.mod.2 %>% 
      dplyr::filter(!mod_rmd %in% c("hs6","hs8"))
    
    for(mod in unique(sort(modv$mod_rmd))){
      
      mod.child <- paste0(mod,".Rmd")
      
      mod.v <- c(knitr::knit_child(input = "hs6.Rmd", envir = globalenv()),
                 knitr::knit_child(input = "hs68.Rmd", envir = globalenv()),
                 knitr::knit_child(input = mod.child, envir = globalenv()))
      
    }
    
  } else {
    
    if(mod.check == "31"){
      
      modv <- tmdl.mod.2 %>% 
        dplyr::filter(!mod_rmd %in% c("hs6","hs7","hs8"))
      
      for(mod in unique(sort(modv$mod_rmd))){
        
        mod.child <- paste0(mod,".Rmd")
        
        mod.v <- c(knitr::knit_child(input = "hs6.Rmd", envir = globalenv()),
                   knitr::knit_child(input = "hs67.Rmd", envir = globalenv()),
                   knitr::knit_child(input = "hs68.Rmd", envir = globalenv()),
                   knitr::knit_child(input = mod.child, envir = globalenv()))
        
      }

    } else {
      
      for(mod in unique(sort(tmdl.mod.2$mod_rmd))){
        
        mod.child <- paste0(mod,".Rmd")
        mod.v <- c(mod.v, knitr::knit_child(input = mod.child, envir = globalenv()))
        
      }
    }
  }
}

#
chap6 <- NULL

for(waterbody_name in unique(sort(model.info$`Model Waterbody`))){
  
  chap6 <- c(chap6, knitr::knit_child(input = "chap6.Rmd", envir = globalenv()))
  
}

```

The following models were used in the development of the `r pro.area.tmdls` for the associated water bodies (`r tbls(name = "tmdl.mod", display="cite")`).

`r tmdl.mod.tbl`

Model calibration was completed by DEQ and described in `r pro.area.tmdls`. It is unlikely that adjustments to the existing calibrated model will occur as part of this project. If it is determined at a later date the calibration needs to be updated the model inputs that are expected to be modified to improve model fit are described in this section. An improved model fit is defined as an improvement to the model fit statistics described in Section 7 Model Evaluation and Acceptance and generally represents improved agreement between measured observations and model predictions.

`r paste(mod.v, collapse = "\n")`

`r paste(chap6, collapse = "\n")`

# Model Evaluation and Acceptance

## Model Uncertainty and Sensitivity

Model uncertainty can arise from a number of sources including error associated with measuring field parameters used for model input or calibration, lack of knowledge on the appropriate value to use for model parameters or constants, or an imperfect mathematical formulation in the model of real world physical processes. A model’s sensitivity is the degree to which predictions are affected by changes in a single or multiple input parameters.

In many cases, the major source of uncertainty is due to uncertainty in spatial representation of the river channel and adjacent landcover (e.g., bathymetry, vegetation height and density) from lack of data or simplification, configuration of the boundary conditions (e.g., uncertainty in estimation of ungaged tributary flows or temperatures), and uncertainty from limited amount or spatial distribution of observed data used for calibration. These sources of uncertainty are largely unavoidable, but do not invalidate the use of the model for decision purposes. 

During the calibration process, it is good practice to evaluate and minimize uncertainty associated with the model parameters to the greatest extent practical (Beck, 1987; EPA, 2009). During the model calibration process, the responsiveness of the model predictions to various assumptions and rate constants should be evaluated. The model setup should include parameters based on literature recommendations and best professional judgment.

Reducing uncertainty in measured field parameters used for model input and calibration is accomplished in the following ways:

*	Data used for the TMDL must have been collected based on a project plan with quality assurance and quality control protocols for collecting and analyzing samples.    

*	The sampling and laboratory analysis must follow widely accepted scientific methods and protocols. These may include ODEQ-MOMs, USEPA, USGS and Standard Methods for the Examination of Water and Wastewater. All acceptable methods include applicable precision and accuracy checks.    

*	When possible, accuracy and precision should be evaluated using DEQ’s data validation criteria as outlined in (ODEQ, 2013). The TMDL program uses waterbody results that demonstrate a data quality level of A, B, or E with careful review. For continuous temperature data a data quality of A or B corresponds to an absolute accuracy <= 1.0 $^\circ C$ and absolute precision <= 2.0 $^\circ C$. Data of unknown quality lacking audit and prepost accuracy checks may also be used following a careful review where it is determined the results appear reasonable and free of issues based on professional judgment.    

Uncertainties in the mathematical formulation are addressed by using open source models that allow free and transparent inspection of model code; and models that have had their methodologies peer reviewed and evaluated.

The existing [A, B, and C] models were extensively evaluated to determine uncertainty in [X, and Y]. [A sensitivity analysis was also completed on Z]. A discussion of the methodology used to evaluate uncertainty and sensitivity is found in [citation]. **[This paragraph should be manually edited for each QAPP with the relevant details.]**

It is not anticipated that additional uncertainty or sensitivity analyses will be performed on the existing calibrated models. 

## Model Acceptance

This section identifies the model acceptance criteria. Model acceptance relies on satisfying seven (7) conditions:

1)	Incorporation of all available field observations of the system (e.g., geometry, flow, boundary inputs/withdrawals, and meteorology) for the time period simulated.    

2)	Model parameters and unmeasured boundary conditions that are within literature-supported and physically defensible ranges.    

3)	Model predicted results have been compared with the associated observed measurements using graphical presentations. Visual comparisons are useful in evaluating model performance over the appropriate temporal or spatial scales.    

4)	Calibration error statistics have been calculated comparing the model predicted results to the associated observed measurements. The calibration error statistics are shown in Equation 2 through Equation 5.    

5)	Calibration error statistics have been used to inform the appropriate use of the model. Where a model achieves an excellent fit it can generally assume a strong role in evaluating management options. Conversely, where a model achieves only a fair or poor fit it should assume a much less prominent role in the overall weight-of-evidence evaluation of management options. If a desired level of quality is not achieved on some or all measures, the model might still be useful; however, a detailed description of its potential range of applicability will be provided.    

6)	Written documentation of all important elements in the model, including model setup, model parameterization, key assumptions, and known areas of uncertainty.    

7)	Peer review as described in Section 9.    

Equation 2 through Equation 4 are the error statistics to be calculated for each calibrated model.

**Mean Error (ME)**: A mean error of zero indicates a perfect fit. A positive value indicates on average the model predicted values are less than the observed data. A negative value indicates on average the model predicted values are greater than the observed data. The mean error statistic may give a false ideal value of zero (or near zero) if the average of the positive deviations between predictions and observations is about equal to the average of the negative deviations in a data set. Because of this, the mean absolute error (MAE) statistic should be used in conjunction with mean error to evaluate model performance.

$$ME = \frac{1}{n}\sum(X_{mod} - X_{obs}) \hspace{3em}\text{Equation 2}$$

**Mean Absolute Error (MAE)**: A mean absolute error of zero indicates a perfect fit. The magnitude of the mean absolute error indicates the average deviation between model predicted values and observed data. The mean absolute error cannot give a false zero.

$$MAE = \frac{1}{n}\sum|X_{mod} - X_{obs}|  \hspace{3em}\text{Equation 3}$$

**Root Mean Square Error (RMSE)**: A root mean square error of zero indicates a perfect fit. Root mean square error is a measure of the magnitude of the difference between model predicted values and observed data.

$$RMSE = \sqrt{\frac{1}{n}\sum(X_{mod} - X_{obs})^2} \hspace{3em}\text{Equation 4}$$

**Nash-Sutcliffe efficiency coefficient (NS)**: Nash-Sutcliffe efficiencies can range from -$\infty$ to 1. An efficiency of 1 corresponds to a perfect match of modeled predicted value to the observed data. An efficiency of 0 indicates that the model predictions are as accurate as the mean of the observed data, whereas an efficiency less than zero occurs when the observed mean is a better predictor than the model.

$$NS = 1 - \frac {\sum (X_{mod} - X_{obs})^2}{\sum (X_{mod} - \overline{X_{obs}})^2}  \hspace{3em}\text{Equation 5}$$

where,  
$X_{mod}$ = The model predicted results;  
$X_{obs}$ = The observed or measured results;  
$\overline X_{obs}$ = The mean of the observed or measured temperature;  
$n$ = The sample size.

# Documentation in Model Reports

The existing TMDL technical appendices will be cited or added to the replacement TMDL as new appendices if those documents contain information supporting replacement TMDL results or decisions. The existing technical appendices will be amended as necessary to document any modifications, updates, or new models that have been developed.

# Peer Review

Peer review of the models and model results will be conducted in the following ways:

DEQ will conduct internal peer review during the modeling process with input from EPA Region 10 as needed.

DEQ will consider feedback on model scenarios and results from the TMDL advisory group and make changes as appropriate.

DEQ will review and respond to any public comments received on the model and model results, and make changes as appropriate.

# Management Scenarios

Management scenarios described in this section summarize means by which the current conditions and possible alternatives will be evaluated. Some or all of these scenarios will be executed and assessed in the modeling.

## Scenario Definitions

### Current Conditions

This scenario evaluates the stream temperature response under current existing conditions. This scenario is similar to the calibrated model except that some sources conditions will be modified, may be removed, or new ones added to reflect the current conditions or discharge loads if they are significantly different from the calibrated model. Elements of this scenario include:

**[List specific changes that will be made]**

### Background

This scenario evaluates the stream temperature response from background sources only. Background sources include all sources of pollution or pollutants not originating from human activities. Background sources may also include anthropogenic sources of a pollutant that DEQ or another Oregon state agency does not have authority to regulate, such as pollutants emanating from another state, tribal lands, or sources otherwise beyond the jurisdiction of the state (OAR 340-042-0030(1)). This scenario essentially combines the following model scenarios: restored vegetation, restored stream flow, improvements to channel morphology, and potentially elements of the climate scenario. The background scenario will be compared to the current conditions model scenario to determine the point of maximum impact, and the amount of cumulative warming originating from human activities that DEQ or another Oregon state agency have authority to regulate.

### Restored Vegetation

This scenario evaluates the stream temperature response with streamside vegetation at restored conditions. The stream temperature warming or cooling contributed by removal of streamside vegetation is evaluated by comparing this scenario to the current condition model. Elements of this scenario include:

* Restoring streamside vegetation in areas along the model extent that are currently characterized as lacking streamside vegetation because of anthropogenic disturbance . The restored vegetation type, height, density, and overhang are identified in the `r paste0( pro.area.tmdls, " (", knitr::combine_words(strip_alpha(unique(sort(model.info$"Abbreviated Reference")))), ")")`.

*	Model inputs for land cover height, canopy density, overhang will be modified to reflect the restored conditions.

*	All other model inputs will be the same as the current condition model.

### Protected Vegetation

This scenario evaluates the stream temperature response from only streamside vegetation along the stream that is currently protected by statue, rule, ordinance, or some other approved management plan (voluntary or regulatory). The purpose of this scenario is to determine the stream temperature warming or cooling contributed by removal of streamside vegetation in unprotected areas and if existing management strategies achieve surrogate measure effective shade targets. This scenario may be a subset of the TMDL implementation scenario. The stream temperature warming or cooling is evaluated by comparing this scenario to the current condition model. Attainment of the effective shade targets is evaluated by comparing this scenario to the background model scenario.  Elements of this scenario include:

*	Identifying streamside vegetation areas along the model extent that are protected and will not be removed. The exact definition of a protected area will be determined during the TMDL process.    

*	Model inputs for land cover height, density, and overhang will be set to zero outside of protected areas.     

*	Model inputs for land cover height, density, and overhang inside a protected areas will be set to current conditions.    

*	All other model inputs will be the same as the current condition model.    

### Restored Stream Flow

This scenario evaluates stream temperature response by maintaining permitted water withdrawals as instream flow. The stream temperature warming or cooling from keeping permitted water withdrawals as instream flow is evaluated by comparing this scenario to the current conditions model scenario. Elements of this scenario include:

*	Maintaining all currently permitted water withdrawals as instream flow in order to increase the thermal loading capacity and reduce stream warming.    

*	Model boundary and tributary flows will be set to reflect the additional instream flows.    

*	All other model inputs will be the same as the current condition model.    

### Tributary Temperatures

This scenario evaluates the stream temperature response from restoration actions on tributaries. The stream temperature warming or cooling contributed by removal of streamside vegetation on tributaries is evaluated by comparing this scenario to the current condition model. Elements of this scenario include:

*	Tributary inputs will be set to reflect restored temperature and flow conditions. The tributary flow will reflect maintaining all currently permitted water withdrawals as instream flow.    

*	All other model inputs will be the same as the current condition model.    

### Climate

This scenario evaluates the stream temperature response from changes in air temperature and relative humidity connected to human caused changes to global or micro climate conditions. Warming or cooling from climate related impacts will be evaluated by comparing this scenario to the current conditions model scenario. Elements of this scenario include:

*	Model inputs for air temperature and relative humidity may be modified to reflect potential conditions or improved climate change scenarios if enough information exists that would allow downscaling to the site specific conditions in model extent.    

*	Model inputs for groundwater or stream flow may also be modified if sufficient information exists that would allow downscaling to the site specific conditions in model extent.    

*	All other model inputs will be the same as the current condition model.    

### Channel Morphology

This scenario evaluates stream temperature response from improvements to channel morphology, including projects to restore cold water refuges. The warming or cooling from channel morphology improvements is evaluated by comparing this scenario to the current conditions model scenario. Elements of this scenario include:    

*	Modifying channel width and/or depth to reflect locations where improvements to channel morphology are needed. The location of channel morphology projects will be determine during the TMDL process.    

*	Model configurations for bottom width, channel angle z, manning’s n, gradient, elevation, porosity, percent hyporheic flow, hyporheic zone thickness, land cover height, density, and overhang may be modified in areas with improved channel morphology.    

* All other model inputs will be the same as the current condition model.    

### No Point Sources

This scenario evaluates the stream temperature response from removing point source heat load. The stream temperature warming or cooling from permitted NPDES point sources is evaluated by comparing. This scenario to the current conditions model scenario. Elements of this scenario include:    

*	Removal of all point sources from the model.    

*	All other model inputs will be the same as the current condition model.     

### TMDL Waste Load Allocations

This scenario evaluates stream temperature warming or cooling from the TMDL Waste Load Allocations. These scenarios will be compared to the background model scenario to evaluating attainment of the human use allowance allocations. Elements of this scenario include:    

*	Modifying point source discharges to reflect proposed or existing TMDL Waste Load Allocations.    

*	All other model inputs will be the same as the current condition model.    

### TMDL Implementation Plans

This set of scenarios evaluate the stream temperature response from proposed or existing DMA and responsible person TMDL implementation plans. These scenarios will be compared to the background model scenario to evaluate attainment of the human use allowance allocations. It is likely that multiple model scenarios will be developed evaluating a single implementation plan or multiple implementation plans together. Elements of this scenario include:    

*	Modifying streamside vegetation, instream flow, and/or channel morphology to reflect the proposed or existing implementation plan. Translating the plan elements to the modeled landscape conditions will be determined during the TMDL process.    

*	Model inputs for land cover height, density, overhang, boundary condition flow and temperature, bottom width, channel angle z, manning’s n, porosity, percent hyporheic flow, and hyporheic zone thickness, may be modified.    

*	All other model inputs will be the same as the current calibrated model.    

# Project Organization

## Project Team/Roles

```{r, label=`roles`, echo=FALSE}

roles <- readxl::read_xlsx("T:/Temperature_TMDL_Revisions/model_QAPPs/R/data/tables.xlsx",
                           sheet = "roles")

troles <- flextable::regulartable(roles) %>% 
  flextable::set_table_properties(layout = "autofit") %>% 
  #flextable::merge_h(i = c(1,13)) %>% 
  #flextable::bold(i = c(1,13)) %>% 
  flextable::bg(i = 1, bg = "#B52222", part = "body") %>% 
  flextable::valign(valign = "top") %>% 
  flextable::theme_box() %>% 
  flextable::set_caption(tbls(name = "roles",
                              caption = paste0("The roles and responsibilities of each team member involved in the temperature TMDL replacement project.")))

```

Project roles and responsibilities are described in `r tbls(name = "roles", display="cite")`.

`r troles`

## Expertise and Special Training Requirements

Additional expertise or special training is not necessary at this time.

DEQ staff involved in developing and configuring models, performing model calibration, running model scenarios, and analyzing and interpreting model results have experience in these tasks from numerous other modeling projects. The Project Manager has extensive experience managing large complex projects and will ensure strict adherence to the project protocols.


## Reports to Management

The DEQ Project Manager (or designee) will provide monthly progress reports to DEQ Management and USEPA. As appropriate, these reports will provide information on the following:

*	Adherence to project schedule and/or budget
*	Deviations from approved QAPP, as determined from project assessment and oversight activities
*	The impact of any deviations on model application quality and uncertainty
*	The need for and results of response actions to correct any deviations
*	Potential uncertainties in decisions based on model predictions and data
*	Data quality assessment findings regarding model input data and model outputs

## Project Schedule

The project schedule for the **[Name of Project TMDL]** is scheduled to occur in two phases. The pre TMDL project phase, and the TMDL and WQMP development phase.

**Pre TMDL project Phase**

The pre TMDL project phase will generally occur between January 2020 through the end of March 2022. In this phase most of the planning and technical work occurs. Specific tasks include:

**Task P1** Data gathering and project organization.

|       **P1.1** Organize and gather effluent data from all active NPDES permittees in the temperature TMDL replacement project area.     

|       **P1.2** Organize and gather all available and relevant river temperature, stream flow, habitat, effective shade, and channel morphology.     
  
|       **P1.3** Complete an open data solicitation.  During the solicitation period, the public may submit continuous stream temperature data and NPDES effluent data to DEQ in the watersheds subject to the temperature TMDL replacements.     

|       **P1.4** Review data collected. Data submitted to DEQ will be screened for completeness and quality, and whether the results are within the typical range expected for that season and time of day.     

|       **P1.5** Stream temperature data will be made available in DEQ’s Ambient Water Quality Monitoring System database (AWQMS).     
  
**Task P2** Develop modeling Quality Assurance Project Plans (QAPPs).  The modeling QAPPs will identify the available data and overall technical approach to be taken for each TMDL project.     

**Task P3** Mapping of Designated Management Agencies (DMAs) and Responsible Persons for counties that are within the project area. All Oregon counties are within the project area except Tillamook and Clatsop, and Deschutes counties.     

**Task P4** Development of computer code to streamline analysis tasks and TMDL document production.     

**Task P5** Development of template TMDL and WQMP section outline and language.     

**Task P6** Implement Modeling QAPPs. This task a follow-up to Task P2. Gathering of new data and completion of new technical described in the modeling QAPPs.     

**TMDL and WQMP development phase**    

The TMDL and WQMP development phase is scheduled to begin **[begin date of task 1]** with EPA’s final agency action approving or disapproving of the TMDL no later than `r format(as.Date(qapp_project_areas[which(qapp_project_areas$areas == qapp_project_area),]$complete.date,"%m/%d/%Y"),"%B %d, %Y")`. In this phase, the draft TMDL and WQMP documents will be written; a TMDL advisory committee will be convened to discuss the updated TMDL allocations and any revisions to the WQMP; and finally DEQ will conduct a public comment period. DEQ will respond to all public comments received, revise the TMDL and WQMP as necessary, and issue the final TMDL to EPA for their action.

# Data Management

DEQ does not anticipate collecting additional field samples. Water quality data gathered and used for this project will be managed in DEQ’s databases AWQMS database.

The modeling software to be used for this project is available on DEQ’s TMDL program website (https://www.oregon.gov/deq/wq/tmdls/Pages/TMDLs-Tools.aspx). 

Model-generated data resulting from testing, calibration, and scenarios will be stored in spreadsheets and text files by DEQ in TMDL specific.  Metadata describing the content, date, and personnel involved in modeling will be documented alongside raw and summarized data.

Secondary data developed as part of this task will be maintained as hardcopy only, both hardcopy and electronic, or electronic only, depending on their nature. 

All electronic data will be maintained on DEQ’s computers and servers. DEQ’s computers are serviced by in-house specialists. When a problem with DEQ’s computers and servers occurs, in-house computer specialists diagnose the problem and correct it if possible. When outside assistance is necessary, the computer specialists call the appropriate vendor. For other computer equipment requiring outside repair and not covered by a service contract, local computer service companies are used on a time-and-materials basis. 

Routine maintenance of DEQ’s computers and servers is performed by in-house computer specialists. Electric power to each computer flows through a surge suppressor to protect electronic components from potentially damaging voltage spikes. All computer users have been instructed on the importance of routinely archiving work assignment data files from hard drive to server storage. The office network server is backed up on tape nightly during the week. Screening for viruses on electronic files loaded on DEQ’s computers or the network is standard policy. Automated screening systems have been placed on all computer systems and are updated regularly to ensure that viruses are identified and destroyed. Annual maintenance of software is performed to keep up with evolutionary changes in computer storage, media, and programs.


# Recordkeeping and Archiving

All data and documents generated during the course of the TMDL project will be achieved according to the current Oregon State Archives Records Retention Schedule. Generally TMDL documents will be retained until 15 years after the TMDL is no longer operational.

The following information will be included in hard copy or electronic format in the TMDL project and Master TMDL folders located on DEQ’s TMDL server.

The project folder will contain at minimum the following subfolders:

Project Plans: All documents related to project planning, project proposals, project schedules, and the QAPPs. Each will reside in their relevant subfolders. The final versions of documents will be clearly identified from drafts and ideally located in separate folders.

Data: All field data organized or collected in support of the TMDL project. This may include water quality samples, NPDES effluent data, field sheets, photos, monitoring metadata, sampling project plans, QAPPS, or other documentation. The data should be organized by parameter and data source if possible.

Models: All models used for the TMDL project including calibration and scenario models. The models should be organized into subfolders for each model domain and model scenario. Draft models and the final TMDL model will be clearly identified and ideally in separate folders. The model folders should include:

*	The model with all input and outputs files; and any executable code used    

*	Copy of all raw and summarized data (including GIS files) used for model input with data source and location metadata included    

*	Scripts or spreadsheets used to transform raw data or used to derive model inputs    

*	Key assumptions and documentation for the model setup and parameterization    

*	Documentation of newly developed model code or modifications to the existing model    

*	Identification of staff that completed the model.

Meetings: All documents produced for external meetings including agendas, presentations, posters, and meeting handouts. Material for each meeting will be saved in a subfolder organized by date and meeting type. For example the folder name for the first meeting of the TMDL advisory group (AG) would be “2022-08-15 Temperature AG 1”. Draft documents and final documents will be clearly identified and ideally in separate folders.

TMDL documents: At each key stage of TMDL and WQMP development copies of the following documents will be saved in separate subfolders within the project folder on the Master TMDL directory.

*	Briefing Memo to DEQ Water Quality Division Administrator or Director.    

*	Public Comment Draft:     

    +	Public Comment Draft TMDL and WQMP Report (Both Word and PDF)    
    
    +	Draft TMDL Appendices (Both Word and PDF)    
    
    +	Public Notice document    
    
    +	TMDL Summary Fact Sheet    
    
    +	News release    
    
    +	GovDelivery Notice and email    
    
    +	Other public notification emails    
    
    +	Mailing List (if used)    
    
    +	Public Comments Errata    
    
*	Public Comments Received: Copy of all public comments received    

*	Final TMDL and WQMP documents:    

    +	TMDL signed Order (both Word and PDF)    
    
    +	TMDL issuance letter to EPA (both Word and PDF)    
    
    +	EPA approval letter (EPA)    
    
    +	Response to Comment Document (both Word and PDF)    
    
    +	TMDL and WQMP Report (both Word and PDF)    
    
    +	TMDL Appendices (both Word and PDF)    
    
    +	TMDL Summary Fact Sheet    
    
    +	News release    
    
    +	GovDelivery Notice and email    
    
    +	Other public notification emails    
    
    +	DMA/RP notification letters (both Word and PDF)    
    
    +	Addendums    
    
    +	Errata    
    
    +	Petitions    
    
    +	Director’s Petition Action (acceptance or rejection of petition)    
    
    +	Response to Petition    
    
    +	Attains upload files    

# QAPP Review and Approval

The DEQ Project Technical Lead will distribute the draft QAPP to the respective DEQ and EPA project team members for review. Comments will be provided to the Project Technical Lead for further discussion if appropriate, and revision and submittal of the final plan within 10 business days of receipt of comments. Following approval, the Project Technical Lead will distribute the final, signed copy to their respective staff assigned to the project. Official copies of the final, approved QAPP will be retained. If any change(s) in the QAPP are required during the project, they must be described in a memorandum and approved by the signatories to this QAPP and attached to the QAPP.

# Implementation and Adaptive Management

```{r, label=`risk`, echo=FALSE}

risks <- readxl::read_xlsx("T:/Temperature_TMDL_Revisions/model_QAPPs/R/data/tables.xlsx",
                           sheet = "risks")

trisks <- knitr::kable(risks, format = "pandoc", padding = 2,
                       caption = tbls(name = "risks",
                                      caption = paste0("Projects risks and proposed solutions.")))

```

DEQ plans to develop a Risk Management Plan to identify project constraints, the risks that may arise during project implementation, and potential solutions. Identified project constraints include the abbreviated project schedule with hard deadlines establish via court order, limited resources, uncertain funding from EPA, and a complex TMDL technical effort which may require additional time and public process. Projects risks from these constraints and proposed solutions are described in `r tbls(name = "risks", display="cite")`. 

`r trisks`

Should a situation arise that requires a significant change in the technical approach, the project team will update the QAPP as needed through revisions or addenda.

# References

```{r, label=`qapp_ref`, echo=FALSE}

tmdl.ref <- model.info %>% 
  #dplyr::filter(`QAPP Project Area` %in% qapp_project_area) %>% 
  dplyr::distinct(`TMDL Document`, `Abbreviated Reference`) %>% 
  dplyr::left_join(ref, by = "Abbreviated Reference") %>% 
  dplyr::mutate(`Abbreviated Reference` = strip_alpha(`Abbreviated Reference`)) %>% 
  dplyr::mutate(reference = paste(`Full Reference`, ifelse(is.na(`Reference Link`),"",`Reference Link`))) %>% 
  dplyr::select(reference)

tir.ref <- data.frame(unique(sort(t52$Reference))) %>% 
  dplyr::rename(reference = "unique.sort.t52.Reference..")

mod.ref <- model.info %>% 
  dplyr::filter(!is.na(mod_ref)) %>% 
  dplyr::rename(reference = mod_ref) %>% 
  dplyr::select(reference)
  

other.ref <- data.frame(reference = c('Beck, M.B. 1987. "Water Quality Modeling: A Review of the Analysis of Uncertainty". Water Resources Research 23(8), 1393',
                                      'EPA (U.S. Environmental Protection Agency). 2009. "Guidance on the Development, Evaluation, and Application of Environmental Models". Council for Regulatory Environmental Modeling, Washington D.C., EPA/100/K-09/003.',
                                      'ODEQ (Oregon Department of Environmental Quality). 2013. "Data validation criteria for water quality parameters measured in the field. DEQ04-LAB-0003-QAG Version5.0." http://www.oregon.gov/deq/FilterDocs/DataQualMatrix.pdf'))


qapp.ref <- rbind(tmdl.ref,tir.ref,mod.ref,other.ref) %>% 
  dplyr::arrange(reference)

for(r in unique(sort(qapp.ref$reference))) {
  
  cat(paste0(r,"    "))
  cat("\
  
      ")
  cat("\n")
  
}

```

