---
output: 
  word_document:
    reference_docx: T:/Temperature_TMDL_Revisions/model_QAPPs/R/data/Report_Template.docx
    keep_md: yes
    toc: yes
    fig_caption: yes
    fig_width: 12
    fig_height: 6
---

```{r setup_hs6, include=FALSE}

knitr::opts_chunk$set(results = "asis",
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      autodep = TRUE)

options(kableExtra.auto_format = F)

```

```{r, "shadow", echo=FALSE}

shadow <- model.info %>% 
  dplyr::filter(`Model version` == "SHADOW") %>% 
  dplyr::select(`Model Waterbody`,`Model Extent Distance`,`Model Extent Units`,`TMDL Document`) %>%
  dplyr::arrange(as.numeric(`Model Extent Distance`))
  

names(shadow)[names(shadow) == "Model Extent Distance"] <- if(NROW(shadow)>0){paste0("Distance (",unique(shadow$`Model Extent Units`),")")}

shadow_tbl <-  knitr::kable(shadow[,c(1:2)], format = "pandoc", padding = 2,
                            caption = tbls(name = "shadow",
                                           caption = paste0("SHADOW model waterbodies in the ", qapp_project_area,".")))
```

### SHADOW

The SHADOW model version 2.3 is a shade, solar input and stream temperature model developed by the U.S. Forest Service (USFS, 1993). 

## SHADOW

DEQ used USFS SHADOW modeling results in its development for the original `r knitr::combine_words(unique(sort(shadow$'TMDL Document')))`. DEQ does not intend to remodel any of the stream segments for the `r qapp_project_area`.  If the need arises to revisit some of the SHADOW modeling, this QAPP will be updated at that time to inform the project team on how to proceed. Complete documentation of the model methodology and inputs are discussed in the SHADOW user's manual (USFS, 1993).

### Model Domain

The SHADOW model domain includes all of the streams identified in `r tbls(name = "shadow", display="cite")`. Typically the extent is the entire length of the stream from mouth to headwaters.

`r shadow_tbl`